### 2021-09-13 proposal/project-proposal.tex
Changing File Name

%%
%% This is file `sample-sigplan.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigplan')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigplan.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigplan,nonacm]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
	\providecommand\BibTeX{{%
			\normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
%%\setcopyright{acmcopyright}
\setcopyright{none}
%%\copyrightyear{2018}
%%\acmYear{2018}
%%\acmDOI{10.1145/1122445.1122456}

%% These commands are for a PROCEEDINGS abstract or paper.
%%\acmConference[Woodstock '18]{Woodstock '18: ACM Symposium%% on Neural
%%  Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
%%\acmPrice{15.00}
%%\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{CS59200 Research Proposal}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Ruiqi Gao}
\email{gao606@purdue.edu}
\affiliation{%
  \institution{Purdue University}
  \city{Lafayette}
  \state{Indiana}
  \country{USA}
  \postcode{47901}
}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
%%\renewcommand{\shortauthors}{Trovato and Tobin, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Various approaches have been developed like data parallelism, model parallelism, attribute parallelism, and pipeline parallelism to train machine learning models in parallel, however, there are no existing frameworks that can integrate these approaches together. Most commonly used machine learning frameworks such as TensorFlow and Pytorch only support a very small subset of parallel paradigms - the data parallelism and model parallelism. GPipe and PipeDream support different types of pipeline parallelism by dividing the model into several modules and run them in a pipeline fashion.
AIRCop(Abstract IR for Comprehensive Parallelization) is a developing project that unifies all the parallelism paradigms mentioned above. Since all the machine learning models can be viewed as computation graphs, the parallelism paradigms (data parallelism, model parallelism, attribute parallelism, and pipeline parallelism) are just different kinds of computation graph transformation, ie transformation pass. AIRCop uses this idea and takes a machine learning model(basic sequential model) and parallelization annotations (describe the desired parallel behavior) as input and generates target program through different transformation passes.

\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Parallel Training, Graph Transformation, Machine Learning Frameworks, Pipeline Parallelism}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Objective}
I am trying to implement pipeline parallelism into a Parallelism Machine learning framework-AIRCop. If the project is done, AIRCop will be the first Machine learning framework that can integrate all parallel paradigms and users can have a deeper insight into parallel training design.

\section{Current Work}
There are currently several parallel strategies to run large models: Spatial Partition(data parallelism, model parallelism) and Pipelined Execution. Spatial Partition divides the computation of machine learning operations among different devices and runs them simultaneously. Pipelined Execution runs multiple batches of data in a pipelined fashion.\par
 Spatial Partition is implemented in well-known machine learning frameworks like Tensorflow and Pytorch. GPipe and PipeDream support different kinds of pipeline parallelism. While Spatial Partition enables the model to be trained among multiple devices simultaneously each with a fraction of data, It will require a large number of device communication between operations and will result in inter-device tensor communication overhead. This turns out to be the bottleneck for parallel training on GPU clusters (group of devices) where inter-cluster communication does not have high bandwidth. Pipeline Execution is a better strategy for parallel training among clusters because it divides the model into several modules, runs different modules simultaneously in a pipelined fashion, and only requires tensor communication between modules. \par 
 It will be ideal for large GPU clusters to use Spatial Partition inside the cluster(where devices have high-bandwidth connections) and Pipeline Execution between different clusters(with little communication overhead). However, there is currently no machine learning framework that can integrate both Spatial Partition and Pipeline Execution. 
\section{Innovation}
My project focuses on implementing pipeline parallelism on the AIRCop framework, thus enables AIRCop to be the first framework that can explore the full power of different Parallelism Paradigms. \par 
 AIRCop also develops a formal language to describe the model as a computation graph, and the corresponding transformation procedure for generating the target parallel program for different parallelism paradigms. AIRCop comes with an annotation system for the user to specify what kind of Parallelism Paradigms they want to use: Spatial Partition annotation for operations and Pipelined Execution annotation for modules. Users can explore comprehensive parallel training by adding different annotations to the source model.
 \section{Implementation and Risk}
 The key idea to implement pipeline parallelism in AIRCop is to realize the corresponding computation graph transformation pass for different pipeline strategies and come up with an annotation system to represent different pipeline strategies.\par
 The risk should be low because GPipe and Pipedream already separately implement different pipeline strategies as libraries and Spatial Partition is already implemented in AIRCop as a transformation pass.
\section{Examination and Success metric}
The first step is to formally describe stacked pipeline(GPipe's strategy) and queued pipeline(PipeDream's strategy) as transformation passes in the formal language of AIRCop, and decides the annotation system for the user to specify the strategy in the source program. \par
The mid-term exam will be to implement stacked pipeline inside LMS compiler as a transformation pass for single pipeline parallelism. \par
The final exam will be to implement both pipeline parallelisms, run evaluation benchmarks on motivating examples by different combinations of parallel strategies.

\end{document}
\endinput
%%
%% End of file `sample-sigplan.tex'.

### 2021-09-14 proposal/project-proposal.tex
Add Risk assessment and Step Timeline

@@
  The risk should be low because GPipe and Pipedream already separately implement different pipeline strategies as libraries and Spatial Partition is already implemented in AIRCop as a transformation pass.
+Because the choice of pipeline strategy is giving by the user using annotations, We only need to transform the source sequential model by changing the training loop of different modules and adding communication between modules while changing tensor's scope.
+Thus the risk should be low.
+However, it would still be possible that we might see no performance gains by combining Pipeline Execution and Spatial Partition than only using Spatial Partition.
+Anyway, integrating Pipeline Execution and Spatial Partition will give user flexibility to try different combinations of parallel strategies under different scenarios on a unified framework.
+This should lead to a deeper insight of parallel training.
 \section{Examination and Success metric}
-The first step is to formally describe stacked pipeline(GPipe's strategy) and queued pipeline(PipeDream's strategy) as transformation passes in the formal language of AIRCop, and decides the annotation system for the user to specify the strategy in the source program.
-\par
-The mid-term exam will be to implement stacked pipeline inside LMS compiler as a transformation pass for single pipeline parallelism. \par
-The final exam will be to implement both pipeline parallelisms, run evaluation benchmarks on motivating examples by different combinations of parallel strategies.
+\begin{itemize}
+  \item The first step is to formally describe stacked pipeline(GPipe's strategy) and queued pipeline(PipeDream's strategy) as transformation passes in the formal language of AIRCop, and decides the annotation system for the user to specify the strategy in the source program.
+By the mean time, I would need to get familiar with LMS framework and study the Spatial Partition Implementation in AIRCop.
+This should be done by 9/30.
+  \item The mid-term exam will be to implement stacked pipeline inside LMS compiler as a transformation pass for single pipeline parallelism. This should be done by 10/30.
+  \item The final exam will be to implement both pipeline parallelisms, run evaluation benchmarks on motivating examples by different combinations of parallel strategies. This should be done by 11/30.
+  \item The final report writing and presentation preparation will be done in the remaining time of the semester. 
+\end{itemize}
 

### 2021-09-28 proposal/proposal.tex
Standardize the directory layout, fix format and grammar issues

--- 2021-09-14 proposal/project-proposal.tex
@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[sigplan,nonacm]{acmart}
+%%\documentclass[sigplan,nonacm]{acmart}
+\documentclass[sigplan]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 
@@
 \setcopyright{none}
+\acmJournal{PACMPL}
+\acmVolume{1}
+\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
+\acmArticle{1}
+\acmYear{2021}
+\acmMonth{9}
+\acmDOI{}
 %%\copyrightyear{2018}
@@
 \begin{document}
+\sloppy
 
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{CS59200 Research Proposal}
+\title{Parallel training by Pipeline Execution In AIRCoP}
 
@@
 GPipe and PipeDream support different types of pipeline parallelism by dividing the model into several modules and run them in a pipeline fashion.
-AIRCop(Abstract IR for Comprehensive Parallelization) is a developing project that unifies all the parallelism paradigms mentioned above.
+AIRCop (Abstract IR for Comprehensive Parallelization) is a developing project that unifies all the parallelism paradigms mentioned above.
 Since all the machine learning models can be viewed as computation graphs, the parallelism paradigms (data parallelism, model parallelism, attribute parallelism, and pipeline parallelism) are just different kinds of computation graph transformation, ie transformation pass.
-AIRCop uses this idea and takes a machine learning model(basic sequential model) and parallelization annotations (describe the desired parallel behavior) as input and generates target program through different transformation passes.
+AIRCop uses this idea and takes a machine learning model (basic sequential model) and parallelization annotations (describe the desired parallel behavior) as input and generates target program through different transformation passes.
 
@@
 \section{Objective}
-I am trying to implement pipeline parallelism into a Parallelism Machine learning framework-AIRCop.
+We are going to implement pipeline parallelism into a Parallelism Machine learning framework-AIRCop.
 If the project is done, AIRCop will be the first Machine learning framework that can integrate all parallel paradigms and users can have a deeper insight into parallel training design.
@@
 \section{Current Work}
-There are currently several parallel strategies to run large models: Spatial Partition(data parallelism, model parallelism) and Pipelined Execution.
+There are currently several parallel strategies to run large models: Spatial Partition (data parallelism, model parallelism) and Pipelined Execution.
 Spatial Partition divides the computation of machine learning operations among different devices and runs them simultaneously.
@@
 \par
- It will be ideal for large GPU clusters to use Spatial Partition inside the cluster(where devices have high-bandwidth connections) and Pipeline Execution between different clusters(with little communication overhead).
+ It will be ideal for large GPU clusters to use Spatial Partition inside the cluster (where devices have high-bandwidth connections) and Pipeline Execution between different clusters (with little communication overhead).
 However, there is currently no machine learning framework that can integrate both Spatial Partition and Pipeline Execution.
@@
 However, it would still be possible that we might see no performance gains by combining Pipeline Execution and Spatial Partition than only using Spatial Partition.
-Anyway, integrating Pipeline Execution and Spatial Partition will give user flexibility to try different combinations of parallel strategies under different scenarios on a unified framework.
+Anyway, integrating Pipeline Execution and Spatial Partition will give users some flexibility to try different combinations of parallel strategies under different scenarios on a unified framework.
 This should lead to a deeper insight of parallel training.
@@
 \begin{itemize}
-  \item The first step is to formally describe stacked pipeline(GPipe's strategy) and queued pipeline(PipeDream's strategy) as transformation passes in the formal language of AIRCop, and decides the annotation system for the user to specify the strategy in the source program.
-By the mean time, I would need to get familiar with LMS framework and study the Spatial Partition Implementation in AIRCop.
+  \item The first step is to formally describe stacked pipeline (GPipe's strategy) and queued pipeline (PipeDream's strategy) as transformation passes in the formal language of AIRCop, and decides the annotation system for the user to specify the strategy in the source program.
+I would need to get familiar with LMS framework and study the Spatial Partition Implementation in AIRCop.
 This should be done by 9/30.
@@
 \endinput
-%%
-%% End of file `sample-sigplan.tex'.

### 2021-09-28 report/paper.tex
Standardize the directory layout, fix format and grammar issues

--- 2021-09-13 proposal/project-proposal.tex
@@
-%%
+%%%
 %% This is file `sample-sigplan.tex',
@@
 
-%%
-%% By default, the full list of authors will be used in the page
-%% headers. Often, this list is too long, and will overlap
-%% other information printed in the page headers. This command allows
-%% the author to define a more concise list
-%% of authors' names for this purpose.
-%%\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
-
-%%
-%% The abstract is a short summary of the work to be presented in the
-%% article.
-\begin{abstract}
-Various approaches have been developed like data parallelism, model parallelism, attribute parallelism, and pipeline parallelism to train machine learning models in parallel, however, there are no existing frameworks that can integrate these approaches together.
-Most commonly used machine learning frameworks such as TensorFlow and Pytorch only support a very small subset of parallel paradigms - the data parallelism and model parallelism.
-GPipe and PipeDream support different types of pipeline parallelism by dividing the model into several modules and run them in a pipeline fashion.
-AIRCop(Abstract IR for Comprehensive Parallelization) is a developing project that unifies all the parallelism paradigms mentioned above.
-Since all the machine learning models can be viewed as computation graphs, the parallelism paradigms (data parallelism, model parallelism, attribute parallelism, and pipeline parallelism) are just different kinds of computation graph transformation, ie transformation pass.
-AIRCop uses this idea and takes a machine learning model(basic sequential model) and parallelization annotations (describe the desired parallel behavior) as input and generates target program through different transformation passes.
-
-\end{abstract}
-
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-
-%%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-\keywords{Parallel Training, Graph Transformation, Machine Learning Frameworks, Pipeline Parallelism}
-
-%%
-%% This command processes the author and affiliation and title
-%% information and builds the first part of the formatted document.
 \maketitle
-
-\section{Objective}
-I am trying to implement pipeline parallelism into a Parallelism Machine learning framework-AIRCop.
-If the project is done, AIRCop will be the first Machine learning framework that can integrate all parallel paradigms and users can have a deeper insight into parallel training design.
-
-\section{Current Work}
-There are currently several parallel strategies to run large models: Spatial Partition(data parallelism, model parallelism) and Pipelined Execution.
-Spatial Partition divides the computation of machine learning operations among different devices and runs them simultaneously.
-Pipelined Execution runs multiple batches of data in a pipelined fashion.\par
- Spatial Partition is implemented in well-known machine learning frameworks like Tensorflow and Pytorch.
-GPipe and PipeDream support different kinds of pipeline parallelism.
-While Spatial Partition enables the model to be trained among multiple devices simultaneously each with a fraction of data, It will require a large number of device communication between operations and will result in inter-device tensor communication overhead.
-This turns out to be the bottleneck for parallel training on GPU clusters (group of devices) where inter-cluster communication does not have high bandwidth.
-Pipeline Execution is a better strategy for parallel training among clusters because it divides the model into several modules, runs different modules simultaneously in a pipelined fashion, and only requires tensor communication between modules.
-\par
- It will be ideal for large GPU clusters to use Spatial Partition inside the cluster(where devices have high-bandwidth connections) and Pipeline Execution between different clusters(with little communication overhead).
-However, there is currently no machine learning framework that can integrate both Spatial Partition and Pipeline Execution.
-\section{Innovation}
-My project focuses on implementing pipeline parallelism on the AIRCop framework, thus enables AIRCop to be the first framework that can explore the full power of different Parallelism Paradigms. \par 
- AIRCop also develops a formal language to describe the model as a computation graph, and the corresponding transformation procedure for generating the target parallel program for different parallelism paradigms.
-AIRCop comes with an annotation system for the user to specify what kind of Parallelism Paradigms they want to use: Spatial Partition annotation for operations and Pipelined Execution annotation for modules.
-Users can explore comprehensive parallel training by adding different annotations to the source model.
- \section{Implementation and Risk}
- The key idea to implement pipeline parallelism in AIRCop is to realize the corresponding computation graph transformation pass for different pipeline strategies and come up with an annotation system to represent different pipeline strategies.\par
- The risk should be low because GPipe and Pipedream already separately implement different pipeline strategies as libraries and Spatial Partition is already implemented in AIRCop as a transformation pass.
-\section{Examination and Success metric}
-The first step is to formally describe stacked pipeline(GPipe's strategy) and queued pipeline(PipeDream's strategy) as transformation passes in the formal language of AIRCop, and decides the annotation system for the user to specify the strategy in the source program.
-\par
-The mid-term exam will be to implement stacked pipeline inside LMS compiler as a transformation pass for single pipeline parallelism. \par
-The final exam will be to implement both pipeline parallelisms, run evaluation benchmarks on motivating examples by different combinations of parallel strategies.
 

### 2021-12-11 report/paper.tex
first draft of report paper finished

@@
 
+\graphicspath{{pictures/}}
 %%
@@
 
+\begin{abstract}
+  Various approaches have been developed like data parallelism, model parallelism, attribute parallelism, and pipeline parallelism to train machine learning models in parallel, however, there are no existing frameworks that can integrate these approaches together.
+Most commonly used machine learning frameworks such as TensorFlow and Pytorch only support a very small subset of parallel paradigms - the data parallelism and model parallelism.
+GPipe and PipeDream support different types of pipeline parallelism by dividing the model into several modules and run them in a pipeline fashion.
+  AIRCop (Abstract IR for Comprehensive Parallelization) is a developing project that unifies all the parallelism paradigms mentioned above.
+Since all the machine learning models can be viewed as computation graphs, the parallelism paradigms (data parallelism, model parallelism, attribute parallelism, and pipeline parallelism) are just different kinds of computation graph transformation, ie transformation pass.
+AIRCop uses this idea and takes a machine learning model (basic sequential model) and parallelization annotations (describe the desired parallel behavior) as input and generates target program through different transformation passes.
+
+\end{abstract}
+
+\keywords{Parallel Training, Graph Transformation, Machine Learning Frameworks, Pipeline Parallelism}
+
 \maketitle
 
+\section{Introduction}
+Machine learning is widely used in various fields like data analysis, automatic driving, Natural language processing, computer vision, and it has bring convenience to our every day life.
+Current machine learning model uses neural network for training and inference, and it is not easy to write machine learning code from scratch.
+With the effort from both the industry and academia, several machine learning framework like Pytorch\cite{paszke2019pytorch} and TensorFlow\cite{abadi2016tensorflow} has been proposed to speed up the development of various Machine learning model.\par
+Neural network models use Gradient Decent as an optimizing approach to achieve the optimal result for reducing the prediction error, and it is extremely exhausting to manually implement the differentiation process.
+As a result automatic differentiation\cite{baydin2018automatic} has been adopted to minimize the engineering effort of developing machine learning algorithm.
+And this technique is widely adopted in modern machine learning frameworks.
+\par
+ Current workflow of a machine learning algorithm are typically composed of 3 steps: the forward pass, the Backward pass, and the optimization pass.
+The forward pass is the procedure to calculate the loss (the prediction error).
+The backward pass is the procedure to derive the gradient of weight.
+The optimization pass is the procedure to update the weight using gradient decent.
+And the forward, backward, optimization pass iterates on fractions (batches) of the data.
+Then a machine learning process can be simply viewed as a training loop composed by input tensors, trainable weight tensors and tensor operations.
+\par
+ \begin{figure}[htbp]
+  \centering
+  \includegraphics[scale=0.8]{ParallelDimension}
+  \caption{Different Spatial Partition Parallelism Strategies}
+  \label{fig:paralleldimension}
+\end{figure}
+ Real word machine learning applications are usually very large, they require tremendous amount of data for training and can not work on a single device.
+As a result, various parallel strategies has been proposed for running large models in parallel among multiple distributed devices to achieve optimal performance\ref{fig:paralleldimension}.
+There are different kinds of parallelism strategies, the most common one is Data parallelism\cite{krizhevsky2012imagenet}, where we split input across multiple devices and maintain the same network weight.
+Data parallelism is adopted and developed by most modern machine learning framework (TensorFlow\cite{abadi2016tensorflow}, Pytorch\cite{paszke2019pytorch}).
+Data parallelism requires network weight synchronizations between devices at the end of each iteration, therefore Data parallelism is ideal for computation intensive network with few weight parameters (convolution and pooling layer) but will introduce significant communication overhead for network with large number of weight parameters (fully-connected layer).
+Model parallelism\cite{dean2012large} is achieved by splitting the network weight tensor across devices while broadcast the input tensor.
+Model parallelism reduces the communication overhead of Data Parallelism but is rarely used because there are often dependencies between network layers.
+Attribute parallelism\cite{jia2019beyond} can partition the attribute dimension in both input and weight parameter and require collective communication like AllReduce between operations.
+Experienced developer can combine different Parallelism strategies to achieve optimal performance.
+For example, \cite{krizhevsky2014one} uses data parallelism for convolution layer and model parallelism for fully-connected layer.
+All the above parallelism strategies can be categorized as Spatial Partition because they are based on partition on data/weight/operation.
+\par
+   Different from Spatial Partition where input data/weight is distributed across multiple devices, there is another parallelism strategy: Pipeline Parallelism.
+Pipeline Parallelism divides a machine learning model into several modules with data dependency between them, divides a batch into several micro batches and runs them in parallel in a pipelined fashion.
+Both GPipe\cite{huang2019gpipe} and Pipedream\cite{narayanan2019pipedream} studied and developed Pipeline parallelism.
+Pipeline parallelism is especially useful when the training model is too big to fit inside a single device and there are dependencies between different layer.
+\par
+ As shown in \cite{krizhevsky2014one}, combining different parallelism strategies can achieve better performance than only using one strategy.
+However there is no unifying machine learning framework that can support all the parallelism strategies.
+TensorFlow\cite{abadi2016tensorflow} and Pytorch\cite{paszke2019pytorch} only supports Data Parallelism while GPipe\cite{huang2019gpipe} and Pipedream\cite{narayanan2019pipedream} supports Pipeline parallelism.
+User are limited to only a subset of Parallelism strategies, and can not fully utilize the distributed training system.
+Different distributed system topologies also would require different strategies for optimal performance.
+As we mentioned before, Spatial Partition may introduce frequent tensor communication overhead between operations.
+This overhead maybe negligible for a GPU Cluster with high-bandwith inter connections but will definitely cause significant performance drop for large distributed systems consist of several nodes of GPU cluster with low bandwith connection between them.
+Because high bandwith inter-GPU connection technique like NV-Link can only scale to a limited number of devices, large training server are often deployed across mutiple nodes.
+It will be ideal to divide the model into several modules and assign each module into a cluster, use Spatial partition inside a GPU cluster and adopt pipeline parallelism for modules across different clusters.
+In conclusion, a unifying framework that supports all the Parallelism strategies will give user more flexibility and opportunity to achieving optimal performance for training large-scale network across distributed systems for different models and system topologies, and give developer deeper insight into parallel training design.\par
+ AIRCoP \cite{aircopfeiwang} is the first framework that supports all the parallel parallelisms above thus allows users to fully explore different parallel settings.
+AIRCoP implemented spatial partition before and our work implemented Pipeline parallelism for AIRCoP, thus make AIRCoP the first unifying parallel training framework.\par
+ \begin{figure}[htbp]
+  \centering
+  \includegraphics[scale=0.7]{FormalSyntax}
+  \caption{Formal Syntax of AIRCoP}
+  \label{fig:sample}
+\end{figure}
+ AIRCoP developed a formal syntax to represent tensor computation graph\ref{fig:sample}.
+The syntax mainly has 3 components: Tensor and corresponding TensorType to represent input tensor and weight, Operation and corresponding Operation Type to represent different Machine learning operations like Conv/MatMul/Relu and their input/output Type, Module to represent different modules.
+Operation and Modules are connected through let-bindings.
+And we use annotation of the formal syntax as an approach to formally represent different parallelism strategies.
+We developed annotation systems for both Tensor, Operation and Modules.
+Tensor and Operation annotation represent Spatial Partition by indicating the dimension to partition.
+Module annotation indicates how many micro-batches the mini batch is split into.
+Data parallelism can be achieved through annotate the input tensor indicating which dimension to split.
+Model parallelism can be achieved by annotating the weight tensor.
+Attribute parallelism can be achieved by a combination  of tensor annotation and operation annotation.
+And pipeline parallelism can be simply done by annotation the Modules.
+User only need to write the sequential forward pass of the Machine Learning model and the correct final training loop will be automatically computed through a series of transformation.
+ \section{Pipeline Parallelism}
+ Pipeline Parallelism is ideal for distributed training workload across several nodes with weak inter-connection.
+Simply using Spatial Partition across different nodes will cause significant communication overhead between each operation.
+By splitting the training models into several modules and assigning each modules to different nodes will limit the inter-node communication to the boundary of each module, therefore significantly reduce the inter-node communication overhead.\par
+ \begin{figure}[htbp]
+  \centering
+  \includegraphics[scale=0.5]{NormalPipeline}
+  \caption{Naive multiple modules}
+  \label{fig:multiplemodule}
+\end{figure}
+ Simply splitting modules will cause large amount of stall time (or bubble time) due to complex dependencies between the sequential modules.
+As shown in\ref{fig:multiplemodule}, if we naively spilt the model into 4 modules $m_0,m_1,m_2,m_3$.
+There will be sequential dependencies between the forward pass $F_0, F_1, F_2, F_3$ of modules $m_0,m_1,m_2,m_3$, and reverse dependencies between backward pass $B_3, B_2, B_1, B_0$.
+The dependencies between the forward pass and backward pass will force strict ordering between modules, introduce synchronizations between module boundaries, and cause significant idle time on 4 devices.
+\par
+ \begin{figure}[htbp]
+  \centering
+  \includegraphics[scale=0.7]{stackedpipeline}
+  \caption{Naive multiple modules}
+  \label{fig:stackedpipeline}
+\end{figure}
+ Our approach is similar to GPipe\cite{huang2019gpipe}, we split a batch into multiple micro-batches and run the forward pass and the backward pass of multiple micro-batches sequentially.
+Only the forward pass and backward pass of the same micro batches has dependencies between them.
+Different module of different micro-batches could run in parallel as shown in\ref{fig:stackedpipeline}.
+Pipeline parallelism could allow user to fully utilize distributed devices by minimizing stall time caused by cross-module dependencies.
+\par
+ Pipeline parallelism will insert blocking send/recv operations at the module boundaries to insure necessary execution ordering.
+And the gradient of multiple micro-batches will be accumulated after all the micro-batches finish and updated the weight synchronously for the entire mini-batch.
+Pipeline parallelism can easily scale to large number of devices while maintain ideal utilization of resources.\par
+ Pipeline parallelism will introduce little communication overhead because it will only transfer data at the module boundaries rather than the frequent inter-operation tensor communication introduced by Spatial Partition.
+This feature of Pipeline parallelism makes it suitable not only for high-speed inter-connection system but also weakly connected systems.
+\par
+ Pipeline parallelism in AIRCoP is done through module annotation systems.
+The annotation consists of a integer representing the number of micro-batches (or pipelines).
+User can splitting the model into several modules and annotates them for Pipeline parallelism.
+The resulting training loop is slightly different from naive multiple modules.
+Assume we have $n$ pipelines, the target program will have a top-level training loop for iterating mini-batches.
+Inside each mini-batch iteration, there will be a training loop of size $n$ for iterating the forward pass of the $n$ micro-batches and a training loop of size $n$ for iterating the backward pass of the $n$ micro-batches.
+The gradient will be accumulated inside each backward pass of micro-batches.
+After the backward pass training loop finishes, the weight will be updated based on gradient decent for the entire mini-batch.
+ \section{Implementation}
+ We implement AIRCoP and Pipeline Parallelism in LMS compiler framework\cite{rompf2010lightweight}, and generate CUDA\cite{sanders2010cuda} code with MPI\cite{gropp1999using}/NCCL for communication as the target executable program.\par
+  LMS (Lightweight Modular Staging) is a library-based staging compiler that distinguish stages solely based on Type information.
+LMS is extremely useful for developing DSL through overloading.
+It has a typed frontend making the code generation safe.
+LMS offers lifted (next-stage) version of normal operators and normal control-flow statements makes symbolic code generation easily and convenient.
+AIRCoP is simply a tensor computation DSL embedded inside LMS framework.\par
+ We embedded the formal syntax of AIRCoP inside the IR of LMS.
+LMS has ANF\cite{flanagan1993essence} form sea of Nodes\cite{click1995simple} IR.
+It also has a effect system for dealing with side effects and dependencies introduced by effects.
+LMS can perform Dead Code Elimination and Common Subexpression Elimination, make it sufficient for our tensor computation graph transformation.
+LMS IR is customizable and offers user flexibility to define their own operation while maintain the data and effect dependencies between them.
+We use custom IR nodes to represent Input/Weight Tensor and Machine learning Operations, and using the effect system to represent their dependencies.
+The Module can be simply represented by a Node with a block which contains all of its operations.
+\par
+ The typed frontend source language of AIRCoP only contains the sequential forward computation of the model and annotations.
+The source language does not mutate tensors.
+We use overloading of the typed frontend objects and operations and reflection to generate the corresponding IR of the source program.
+We put the explicit TensorType and Annotations into the rhs of the IR to make it persistent and used in later transformation.
+\par
+ We use transformation pass of LMS to transform the IR and simulate the tensor graph transformation to generate the correct training loop.
+We first need to perform DimName pass to merge the equivalent dimension of different tensors.
+For example, if matrix multiplication happens between matrixes of dimension $[m, n]$ and dimension $[j,k]$, then dimension $n$ and dimension $k$ can be merged.
+Annotations of tensors and operations represent which dimension to partition, so we have to update those annotations as well.
+\par
+The next pass DistributeTensorAIRCoP is the core pass to implement training loop transformation. This pass has 2 phases \textbf{Collection Phase} and \textbf{Generation Phase}.\par
+ The \textbf{Collection Phase} collects the forward operations, the weight nodes and the corresponding backward operation of each node.
+The backward operation is collected in reverse order to maintain the correct semantic of differentiation.
+After Collection Phase, we have the necessary information needed to perform automatic differentiation (AD).
+To generate the correct training loop, we also need to collect lifted tensors.
+As we mentioned before, the forward pass and backward pass of micro-batches are iterated in different loops to achieve pipeline parallelism.
+Therefore, the tensors that are allocated in the forward pass and used in the backward pass will cause error because forward and backward pass are now in different scope.
+Those tensors are called lifted tensors, we have to lift them into the outer scope and maintains multiple versions of them as an array of tensors.
+The forward and backward pass of micro-batches will use different versions of them lifted tensors.
+After collect the lifted tensors, the Collection Phase is finished.\par
+ The \textbf{Generation Phase} generates the correct training loop based on the collected information.
+It will add paired send/recv operations at the boundaries of different modules to maintain the correct data dependencies.
+Inside each module, it will first generates the top-level training loop for mini-batches.
+Then it will generate the forward training loop and backward training loop for micro-batches based on Module Annotations and use the collected backward operations to realize automatic differentiation.
+Inside both the forward and backward pass, the read and write to the lifted tensors will be transformed into read/write to the versioned tensor array.
+The weight gradient will be accumulated in the backward pass and optimized in the end for the entire mini-batch.\par
+ The next pass will perform the Spatial Partition transformation.
+This part is of past work so we will not dig further into this pass.
+The main purpose of this pass is to partition the tensor and operations based on annotations and add the corresponding inter operation tensor communication for synchronization (AllReduce for MatrixMultiplication or Convolution).\par
+ The final pass is to generate the corresponding MPI/NCCL instructions for inter-device communications.
+NCCL (NVIDIA Collective Communications Library) is the communication library introduced by NVIDIA for efficient collective and point to point communication across multiple GPU and multiple Nodes.
+Our target program is CUDA code with MPI/NCCL for communication, we use MPI to set up the environment and the NCCL communicator.
+We will create 2 NCCL communicator, one global communicator for cross-module communication (NCCLSend/Recv) and one local communicator for collective tensor communication (NCCLAllReduce) inside modules.
+We also need to handle tensor communication between modules if they have different partitions which can be naively implemented by gather and scatter.
+We use cuBLAS and cuDNN for matrix computation and convolution network.
+To optimize the memory management, we statically allocate GPU memory before training loop and avoid time-exhausting cudaMalloc/cudaFree inside training loop.
+\par
+ Our generated CUDA code can efficiently represent the correct training loop of machine learning model with both Spatial Partition and Pipeline Parallelism.
+ \section{Evaluation}
+ In this section we show the evaluation benchmarks of various forms of Pipeline parallelism by running our generated CUDA code on 4 GeForce TITAN XP GPUs.
+We did not perform comparison between Spatial Partition and Pipeline parallelism because the ideal network topologies for Pipeline parallelism is a large training system consist of several sever Nodes which we do not currently have.
+The purpose of this benchmark is to validate our graph transformation and gain a deeper insight into the choice of different parallelism strategies.\par
+ We perform our test on a micro benchmark (10 convolution layer network) with different parallelism settings:
+ \begin{itemize}
+   \item Naive 2 module parallelism with each module running in Data parallelism on 2 GPUs
+   \item 2 modules in Pipelined parallelism (4 pipelines) with each module running in Data parallelism on 2 GPUs
+   \item 2 modules in Pipelined parallelism (8 pipelines) with each module running in Data parallelism on 2 GPUs
+   \item 4 modules in Pipelined parallelism (4 pipelines) with each module running in a single GPU
+ \end{itemize}
+ \begin{figure}[htbp]
+  \centering
+  \includegraphics[scale=0.4]{runtime.png}
+  \caption{Data Parallelism vs Pipeline parallelism}
+  \label{fig:runtime}
+\end{figure}
+ All the 4 settings runs in 4 GPUs and the running time of 100 iterations on these settings are shown in figure\ref{fig:runtime}.
+We can notice that using Pipelined parallelism rather than naive module partition on 2 modules will cause a speed up of 1.4x.
+And we can further accelerate the training loop by either increase the pipeline number (micro-batch number) or the module number.
+But keep in mind that increase pipeline number will increase memory usage due to versioned tensor and increase module number will increase cross-module communication overhead.\par
+ The speed up achieved may not seem significant but is understandable due to the fact that the communication (NCCLSend/Recv) is blocking and could take considerable time for large data transfer.
+Load in-balancing between different modules could also cause performance drop.\par
+ The evaluation shows that our framework can unify both Spatial Partition and Pipeline Parallelism, validates the correctness of our training loop transformation and offers a deeper insight into how to choose between different parallelism training settings.
+ \section{Conclusion}
+In conclusion, we implement Pipeline parallelism for AIRCoP, make it the first machine learning framework to integrate all the parallelism strategies and offer user flexibility to fully explore different combinations of parallel training strategies under different hardware scenarios and gain a deeper insight into parallel training.
+\bibliographystyle{acm}
+\bibliography{paper}
 \end{document}

### 2021-12-11 report/paper.tex
Second version finished: Add Proposal Retrospect and Progress

--- 2021-09-28 report/paper.tex
@@
 
+\begin{abstract}
+  Various approaches have been developed like data parallelism, model parallelism, attribute parallelism, and pipeline parallelism to train machine learning models in parallel, however, there are no existing frameworks that can integrate these approaches.
+Most commonly used machine learning frameworks such as TensorFlow and Pytorch only support a very small subset of parallel paradigms - the data parallelism and model parallelism.
+GPipe and PipeDream support different types of pipeline parallelism by dividing the model into several modules and running them in a pipeline fashion.
+  AIRCop (Abstract IR for Comprehensive Parallelization) is a developing project that unifies all the parallelism paradigms mentioned above.
+Since all the machine learning models can be viewed as computation graphs, the parallelism paradigms (data parallelism, model parallelism, attribute parallelism, and pipeline parallelism) are just different kinds of computation graph transformation, ie transformation pass.
+AIRCop uses this idea and takes a machine learning model (basic sequential model) and parallelization annotations (describe the desired parallel behavior) as input and generates target program through different transformation passes.
+
+\end{abstract}
+
+\keywords{Parallel Training, Graph Transformation, Machine Learning Frameworks, Pipeline Parallelism}
+
 \maketitle
-
+\section{Proposal Retrospect and Progress}
+\subsection{Proposal Retrospect}
+In the proposal, we mentioned that both stacked pipeline and queued pipeline will be supported by this project. However, we currently only support stacked pipeline for pipeline execution.\par
+The reason is that we spent some time adding support for multiple-module for AIRCoP because AIRCoP originally only support single module.
+This involves changing the syntax of the frontend source language to make writing multiple-module models intuitive and easy.
+After finishing support for the stacked pipeline, We also found a severe problem in AIRCoP that will cause significant performance overhead - Memory Management.
+AIRCoP originally use cudaMalloc and cudaFree inside the training loop for memory allocation.
+Due to the complex synchronization mechanism of cudaMalloc/Free, these operations will cause a very long time.
+After analyzing the running time with NVIDIA profiler, we found that most of the execution time is spent on cudaMalloc/Free and it could be even 10x faster by pre-allocating the memory before the training loop.
+So we focus on addressing the memory management problem after finishing stacked pipeline.
+We currently use some code-motion techniques that will lift all the cudaMalloc out of the training loop for optimization.\par
+Because queued pipeline parallelism uses asynchronous weight update and will use lagged network weight, we choose to address the memory management problem first due to limited time.\par
+Because we only currently implement stacked pipeline, all the pipeline parallelism used in this paper refers to stacked pipeline parallelism.
+\subsection{Progress}
+Although we only support stacked pipeline, it is sufficient for multi-module pipeline parallelism.
+We can still mix spatial partition with pipeline execution to explore different parallelism settings and gain some insight into parallel training as we will see in the evaluation.\par
+Our progress is shown as follows:
+\begin{itemize}
+  \item We formalize the transformation procedure of stacked pipeline as training loop transformation on the formal language of AIRCoP.
+  \item We implemented stacked pipeline for AIRCoP.
+  \item We optimize the memory management for AIRCoP.
+  \item We run evaluation benchmarks with different parallelism settings(Spatial Partition, Spatial Partition mixed with Pipeline Execution) to validate our implementation and gain some insight into parallel training.
+\end{itemize} \par
+In section \ref{introduction}, we introduce parallel training, different existing parallelism strategies, and AIRCoP.\par
+In section \ref{pipelineparallism}, we further study stacked pipeline - the pipeline parallelism we support in AIRCoP.\par
+In section \ref{formalpipe}, we show the formal description of the training loop transformation for pipeline parallelism.\par
+In section \ref{implementation}, we present the implementation detail of our work.\par
+In section \ref{evaluation}, we present a parallel training evaluation with different parallelism settings to illustrate the flexibility of our framework.
+\section{Introduction} \label{introduction}
+ Real-world machine learning applications are usually very large, they require a tremendous amount of data for training and can not work on a single device.
+As a result, various parallel strategies have been proposed for running large models in parallel among multiple distributed devices to achieve optimal performance.\par
+ \subsection{Spatial Partition}There are different kinds of parallelism strategies, the most common one is Data parallelism\cite{krizhevsky2012imagenet}, where we split input across multiple devices and maintain the same network weight.
+Data parallelism is adapted and developed by most modern machine learning framework (TensorFlow\cite{abadi2016tensorflow}, Pytorch\cite{paszke2019pytorch}).
+Data parallelism requires network weight synchronizations between devices at the end of each iteration, therefore Data parallelism is ideal for a computation-intensive network with few weight parameters (convolution and pooling layer) but will introduce significant communication overhead for a network with a large number of weight parameters (fully-connected layer).
+Model parallelism\cite{dean2012large} is achieved by splitting the network weight tensor across devices while broadcasting the input tensor.
+Model parallelism reduces the communication overhead of Data Parallelism but is rarely used because there are often dependencies between network layers.
+Attribute parallelism\cite{jia2019beyond} can partition the attribute dimension in both input and weight parameters and require collective communication like AllReduce between operations.
+An experienced developer can combine different Parallelism strategies to achieve optimal performance.
+For example, \cite{krizhevsky2014one} uses data parallelism for the convolution layer and model parallelism for the fully-connected layer.
+All the above parallelism strategies can be categorized as Spatial Partition because they are based on the partition for data/weight/operation.
+\par
+ \subsection{Pipeline Parallelism}
+ Different from Spatial Partition where input data/weight is distributed across multiple devices, there is another parallelism strategy: Pipeline Parallelism.
+Pipeline Parallelism divides a machine learning model into several modules with data dependency between them, divides a batch into several micro-batches, and runs them in parallel in a pipelined fashion.
+Both GPipe\cite{huang2019gpipe} and Pipedream\cite{narayanan2019pipedream} studied and developed Pipeline parallelism.
+Pipeline parallelism is especially useful when the training model is too big to fit inside a single device and there are dependencies between different layers.
+\par
+ As shown in \cite{krizhevsky2014one}, combining different parallelism strategies can achieve better performance than only using one strategy.
+However, there is no unifying machine learning framework that can support all the parallelism strategies.
+TensorFlow\cite{abadi2016tensorflow} and Pytorch\cite{paszke2019pytorch} only supports Data Parallelism while GPipe\cite{huang2019gpipe} and Pipedream\cite{narayanan2019pipedream} supports Pipeline parallelism.
+Users are limited to only a subset of Parallelism strategies, and can not fully utilize the distributed training system.
+Different distributed system topologies also would require different strategies for optimal performance.
+As we mentioned before, Spatial Partition may introduce frequent tensor communication overhead between operations.
+ \begin{figure}[htbp]
+  \centering
+  \includegraphics[scale=0.7]{networktopo.png}
+  \caption{Large training system}
+  \label{fig:networktopo}
+\end{figure} This overhead may be negligible for a GPU Cluster with high-bandwidth interconnections but will cause a significant performance drop for large distributed systems consisting of several nodes of GPU clusters with low bandwidth connection between them\ref{fig:networktopo}.
+Because high bandwidth inter-GPU connection techniques like NV-Link can only scale to a limited number of devices, large training servers are often deployed across multiple nodes.
+It will be ideal to divide the model into several modules and assign each module into a cluster, use Spatial partition inside a GPU cluster and adopt pipeline parallelism for modules across different clusters.\par
+\subsection{AIRCoP}
+ AIRCoP \cite{aircopfeiwang} is the first framework that supports all the parallel parallelisms above thus allows users to fully explore different parallel settings.
+AIRCoP implemented spatial partition before and our work implemented Pipeline parallelism for AIRCoP, thus make AIRCoP the first unifying parallel training framework.\par
+ %\begin{figure}[htbp]
+ % \centering
+ % \includegraphics[scale=0.7]{FormalSyntax}
+ % \caption{Formal Syntax of AIRCoP}
+ % \label{fig:sample}
+%\end{figure}
+\begin{align*}\label{formalsyntax}
+  \textbf{TensorType}:& \\
+  s \in Symbol& \quad n  \in \mathbb{N} \\
+  size \in Size& :=\mathcal{S}(dim:Symbol,length:\mathbb{N})\\
+  tt\in TensorType& := \mathcal{T}(shape:List[Size])\\
+\end{align*}
+\begin{align*}
+  \textbf{OperationType}:& \\
+  mem\_op\in MemoryOp& := Weight \text{ } | \text{ } Input \text{ } | \text{ } Zero \text{ } | \text{ } One \text{ } \dots \\
+  ml\_op \in MLOP& := MatMul \text{ } | \text{ } Conv \text{ } | \text{ } AvgPool \text{ }\text{ } | \dots \\
+  op \in OperationType&:= mem\_op \text{ } | \text{ } ml\_op \\
+  \textbf{Module}:& \\
+  x \in Var& \quad \quad \quad xs \in List[Var]\\
+  operation \in Operation&:= xs \text{ } | \text{ } op(inputs: List[Var], \\
+  & out\_types:List[TensorType]) \\
+  module \in Module&:=xs \text{ } | \text{ } \textbf{let } xs=operation \textbf{ in } module \\
+  \textbf{Model}:&\\
+  model \in Model&:= \quad List[Module]
+\end{align*}
+ AIRCoP developed a formal syntax to represent tensor computation graphs.
+The syntax mainly has 3 components: Tensor and corresponding TensorType to represent input tensor and weight, Operation, and corresponding Operation Type to represent different Machine learning operations like Conv/MatMul/Relu and their input/output Type, Module to represent different modules.
+Operations and Modules are connected through let-bindings.
+And we use annotation of the formal syntax as an approach to formally represent different parallelism strategies.
+We developed annotation systems for both Tensor, Operation, and Modules.
+Tensor and Operation annotation represents Spatial Partition by indicating the dimension to partition.
+Module annotation indicates how many micro-batches the mini-batch is split into.
+Data parallelism can be achieved through annotating the input tensor indicating which dimension to split.
+Model parallelism can be achieved by annotating the weight tensor.
+Attribute parallelism can be achieved by a combination of tensor annotation and operation annotation.
+And pipeline parallelism can be simply done by annotation the Modules.
+Users only need to write the sequential forward pass of the Machine Learning model and the correct final training loop will be automatically computed through a series of transformations.
+ \section{Pipeline Parallelism} \label{pipelineparallism}
+ Pipeline Parallelism is ideal for distributed training workload across several nodes with weak inter-connection.
+Simply using Spatial Partition across different nodes will cause significant communication overhead between each operation.
+By splitting the training models into several modules and assigning each module to different nodes will limit the inter-node communication to the boundary of each module, therefore significantly reducing the inter-node communication overhead.\par
+ \begin{figure}[htbp]
+  \centering
+  \includegraphics[scale=0.3]{NormalPipeline}
+  \caption{Naive multiple modules}
+  \label{fig:multiplemodule}
+\end{figure}
+ Simply splitting modules will cause large amount of stall time (or bubble time) due to complex dependencies between the sequential modules.
+As shown in\ref{fig:multiplemodule}, if we naively spilt the model into 4 modules $m_0,m_1,m_2,m_3$.
+There will be sequential dependencies between the forward pass $F_0, F_1, F_2, F_3$ of modules $m_0,m_1,m_2,m_3$, and reverse dependencies between backward pass $B_3, B_2, B_1, B_0$.
+The dependencies between the forward pass and backward pass will force strict ordering between modules, introduce synchronizations between module boundaries, and cause significant idle time on 4 devices.
+\par
+ \begin{figure}[htbp]
+  \centering
+  \includegraphics[scale=0.4]{stackedpipeline}
+  \caption{Pipeline Parallelism}
+  \label{fig:stackedpipeline}
+\end{figure}
+ Our approach is similar to GPipe\cite{huang2019gpipe}, we split a batch into multiple micro-batches and run the forward pass and the backward pass of multiple micro-batches sequentially.
+Only the forward pass and backward pass of the same micro-batches have dependencies between them.
+Different modules of different micro-batches could run in parallel as shown in\ref{fig:stackedpipeline}.
+Pipeline parallelism could allow users to fully utilize distributed devices by minimizing stall time caused by cross-module dependencies.
+\par
+ Pipeline parallelism will insert blocking send/recv operations at the module boundaries to insure necessary execution ordering.
+And the gradient of multiple micro-batches will be accumulated after all the micro-batches finish and updated the weight synchronously for the entire mini-batch.
+Pipeline parallelism can easily scale to a large number of devices while maintaining ideal utilization of resources.\par
+ Pipeline parallelism will introduce little communication overhead because it will only transfer data at the module boundaries rather than the frequent inter-operation tensor communication introduced by Spatial Partition.
+This feature of Pipeline parallelism makes it suitable not only for high-speed interconnection systems but also weakly connected systems.
+\par
+ \begin{align*}
+  \textbf{ModuleAnnotation}: \qquad \qquad& \\
+  a \in PipeAnno  := \mathcal{K} (\epsilon : \mathcal{N})\qquad \qquad&
+\end{align*}
+ Pipeline parallelism in AIRCoP is done through module annotation systems.
+The annotation consists of an integer representing the number of micro-batches (or pipelines).
+Users can split the model into several modules and annotate them for Pipeline parallelism.
+The resulting training loop is slightly different from naive multiple modules.
+Assume we have $n$ pipelines, the target program will have a top-level training loop for iterating mini-batches.
+Inside each mini-batch iteration, there will be a training loop of size $n$ for iterating the forward pass of the $n$ micro-batches and a training loop of size $n$ for iterating the backward pass of the $n$ micro-batches.
+The gradient will be accumulated inside each backward pass of micro-batches.
+After the backward pass training loop finishes, the weight will be updated based on gradient descent for the entire mini-batch.
+ \section{Formal description of Pipeline Parallelism Transformation}\label{formalpipe}
+%\begin{align*}
+%  \textbf{ModuleAnnotation}:& \\
+%  a \in PipeAnno & := \mathcal{K} (\epsilon : \mathcal{N}) \\
+%  \textbf{Model}:&\\
+%   s\_model \in Model&:= \quad List[Module \textbf{ by } PipeAnno] \\
+%   \textbf{Transformation}:&\\
+%   Transform&:Model \rightarrow Model \\
+%   Transform (module::modulelist)&\triangleq Transform(module)::Transform(modulelist)\\
+%   Transform (module \in Module)&\triangleq Collection(module)(\emptyset,\emptyset,\emptyset)\\
+%   \textbf{Collection Phase}:&\\
+%   Collection: Module \rightarrow List[Var \times Operation]& \rightarrow List[Var \times Operation] \rightarrow List[Var \times operation] \rightarrow Module \\
+%   Collection(\textbf{let } xs = &Weight([], ots) \textbf{ in } module)(E_w,E_f,E_b)\triangleq \\
+%   Collection&(module)(E_w :: (xs, Weight([], ots), E_f,E_b) \\
+%   Collection(\textbf{let } xs = &Input([], ots) \textbf{ in } module)(E_w,E_f,E_b)\triangleq \\
+%   Collection&(module)(E_w, E_f :: (xs, Input([], ots)),E_b) \\
+%   Collection(\textbf{let } xs = &ml\_op(ins, ots) \textbf{ in } module)(E_w,E_f,E_b)\triangleq \\
+%   Collection(module)(E_w, &E_f :: (xs, ml\_op(ins, ots)),(ins^\prime, ml\_op^\prime(ins::xs::xs^\prime, ins.tt))::E_b) \\
+%   Collection(xs)&(E_w,E_f,E_b) \quad if \text{ (xs is not used in subsequent module)}\triangleq\\
+%   Generation(E_w, &E_f,(xs^\prime, One([], xs.tt))::E_b) \\
+%   Collection(xs)&(E_w,E_f,E_b) \quad if \text{ (xs is used in subsequent module)}\triangleq\\
+%   Generation(E_w, &E_f::([], Send([xs],xs.tt)),(xs^\prime, Recv([], xs.tt))::E_b) \\
+%   \textbf{Generation Phase}:&\\
+%\end{align*}
+Transformation for pipeline parallelism has 2 stages; Collection Phase and Generation Phase.
+The Collection Phase collects the necessary information like forward/backward operation.
+The Generation Phase uses the collected information and the module annotation to generate the training loop.\par
+During Collection Phase we need to collect the weight operations $E_w$, forward operations $E_f$ and backward operations $E_b$.
+For pipeline parallelism, we will add send/recv into $E_f$ and $E_b$ to maintain the data dependencies between modules.
+The corresponding variables of the operation is $xs^w$, $xs^f$, $xs^b$.
+We also need to calculate the lift variables $xs^l = Defined(E_f) \bigcap Used(E_b)$, the lifted variables $xs^l$ are the variables that are defined in the forward pass and used in the backward pass.
+Because we are generating different loops for forward pass and backward pass of micro-batches, we need to lift $xs^l$ into the outer scope and keep multiple versions of those variables for different micro-batches.\par
+The formal description of the generation phase is as follows:
+\begin{align*}
+  \textbf{Generation Phase}: \qquad \qquad& \\
+  G(E_w,E_f,E_b, xs^w,xs^f,xs^b,xs^l,\alpha:\mathcal{K} (\epsilon))& \triangleq \\
+  G_w(E_w)G_l(xs^l)(loop \text{  } i=[0\dots n/\epsilon] &\textbf{ in}\\
+  loop \text{  }j=[0\dots \epsilon]\text{ }G_f(E_f, xs^l) \textbf{ then}&\\
+  loop \text{  }j=[0\dots \epsilon]\text{ }G_b(E_b, xs^f) \textbf{ then}&\\
+  G_o(xs^w) \textbf{ then } xs^w\quad \qquad \qquad&
+\end{align*}
+\begin{align*}
+  G_w&((xs, Weight([], out\_type))::E_w)\triangleq\\
+  &\textbf{let } xs=Weight([], out\_type) \textbf{ in}\\
+  &\textbf{let } xs^\prime=Zero([], out\_type) \textbf{ in}\\
+  &\textbf{let } xs^m=Zero([], out\_type) \textbf{ in}\\
+  &G_w(E_w)\\
+  G_l&(xs::xs^l)\triangleq\\
+  &\textbf{let } xs^\epsilon=Zero^\epsilon([], xs.type) \textbf{ in}\\
+  &G_l(xs^l)\\
+  G_f&((xs, op([input], out\_type))::E_f, xs^l)\triangleq\\
+  &\textbf{let } xs=op([input], out\_type) \textbf{ in}\\
+  &G_f(E^f, xs^l)\\
+  G_f&(\emptyset, xs::xs^l)\triangleq\\
+  &\textbf{let } \phi=Assign([xs, xs^\epsilon], xs.type) \textbf{ in}\\
+  &G_f(\emptyset, xs^l)
+\end{align*}
+\begin{align*}
+  G_b&(E_b, xs::xs^f)\triangleq\\
+  &\textbf{let } xs^\prime=Zero([], out\_type) \textbf{ in}\\
+  &G_b(E^b, xs^f)\\
+  G_b&((xs^\prime, op^\prime([input], out\_type))::E_b, \emptyset)\triangleq\\
+  &\textbf{let } t=op^\prime_l([input], out\_type) \textbf{ in}\\
+  &\textbf{let } \phi=Accumulate([t, xs^\prime], out\_type) \textbf{ in}\\
+  &G_b(E^b, \emptyset)
+\end{align*}
+\begin{align*}
+  G_o&(xs::xs^w)\triangleq\\
+  &\textbf{let } \phi=Optimize([xs,xs^\prime,xs^m], xs.type) \textbf{ in}\\
+  &G_o(xs^w)\\
+  op_l&([input],out\_type) \triangleq op(input^\epsilon, out\_type)
+\end{align*}
+$op_l$ means that the operation will read from the versioned lift variable instead of the originally variable.
+$xs^\epsilon = Zero^\epsilon$ will generates an array of $\epsilon$ versions of the variable $xs$.\par
+$G_w$ generates the weight and their gradient and momentum, $G_l$ generate an array of lift nodes for different micro-batches, $G_f$ generates forward pass and write the correct value to the versioned lift variables, $G_b$ generates the backward pass and read from the versioned lift variable, $G_o$ generates the optimization pass.
+Because our source language does not allow mutation, we only need to write the versioned lift node at the end of the forward pass and the Generation phase can generate the correct training loop.
+ \section{Implementation} \label{implementation}
+ We implement AIRCoP and Pipeline Parallelism in LMS compiler framework\cite{rompf2010lightweight}, and generate CUDA\cite{sanders2010cuda} code with MPI\cite{gropp1999using}/NCCL for communication as the target executable program.\par
+  LMS (Lightweight Modular Staging) is a library-based staging compiler that distinguishes stages solely based on Type information.
+LMS is extremely useful for developing DSL through overloading.
+It has a typed frontend making the code generation safe.
+LMS offers lifted (next-stage) versions of normal operators and normal control-flow statements make symbolic code generation easy and convenient.
+AIRCoP is simply a tensor computation DSL embedded inside LMS framework.\par
+ We embedded the formal syntax of AIRCoP inside the IR of LMS.
+LMS has ANF\cite{flanagan1993essence} form sea of Nodes\cite{click1995simple} IR.
+It also has an effect system for dealing with side effects and dependencies introduced by effects.
+LMS can perform Dead Code Elimination and Common Subexpression Elimination, making it sufficient for our tensor computation graph transformation.
+LMS IR is customizable and offers users the flexibility to define their operation while maintaining the data and effect dependencies between them.
+We use custom IR nodes to represent Input/Weight Tensor and Machine learning Operations and use the effect system to represent their dependencies.
+The Module can be simply represented by a Node with a block that contains all of its operations.
+\par
+ The typed frontend source language of AIRCoP only contains the sequential forward computation of the model and annotations.
+The source language does not mutate tensors.
+We use overloading of the typed frontend objects and operations and reflection to generate the corresponding IR of the source program.
+We put the explicit TensorType and Annotations into the RHS of the IR to make it persistent and used in later transformation.
+\par
+ We use the transformation pass of LMS to transform the IR and simulate the tensor graph transformation to generate the correct training loop.
+We first need to perform DimName pass to merge the equivalent dimension of different tensors.
+For example, if matrix multiplication happens between matrixes of dimension $[m, n]$ and dimension $[j,k]$, then dimension $n$ and dimension $j$ can be merged.
+Annotations of tensors and operations represent which dimension to partition, so we have to update those annotations as well.
+\par
+The next pass DistributeTensorAIRCoP is the core pass to implement training loop transformation. This pass has 2 phases \textbf{Collection Phase} and \textbf{Generation Phase}.\par
+ The \textbf{Collection Phase} collects the forward operations, the weight nodes, and the corresponding backward operation of each node.
+The backward operation is collected in reverse order to maintain the correct semantic of differentiation.
+After Collection Phase, we have the necessary information needed to perform automatic differentiation (AD).
+To generate the correct training loop, we also need to collect lifted tensors.
+As we mentioned before, the forward pass and backward pass of micro-batches are iterated in different loops to achieve pipeline parallelism.
+Therefore, the tensors that are allocated in the forward pass and used in the backward pass will cause an error because forward and backward passes are now in different scopes.
+These tensors are called lifted tensors, we have to lift them into the outer scope and maintain multiple versions of them as an array of tensors.
+The forward and backward pass of micro-batches will use different versions of the lifted tensors.
+After collecting the lifted tensors, the Collection Phase is finished.\par
+ The \textbf{Generation Phase} generates the correct training loop based on the collected information.
+It will add paired send/recv operations at the boundaries of different modules to maintain the correct data dependencies.
+Inside each module, it will first generate the top-level training loop for mini-batches.
+Then it will generate the forward training loop and backward training loop for micro-batches based on Module Annotations and use the collected backward operations to realize automatic differentiation.
+Inside both the forward and backward pass, the read and write to the lifted tensors will be transformed into read/write to the versioned tensor array.
+The weight gradient will be accumulated in the backward pass and optimized in the end for the entire mini-batch.\par
+ The next pass will perform the Spatial Partition transformation.
+This part is of past work so we will not dig further into this pass.
+The main purpose of this pass is to partition the tensor and operations based on annotations and add the corresponding inter operation tensor communication for synchronization (AllReduce for MatrixMultiplication or Convolution).\par
+ The final pass is to generate the corresponding MPI/NCCL instructions for inter-device communications.
+NCCL (NVIDIA Collective Communications Library) is the communication library introduced by NVIDIA for efficient collective and point-to-point communication across multiple GPU and multiple Nodes.
+Our target program is CUDA code with MPI/NCCL for communication, we use MPI to set up the environment and the NCCL communicator.
+We will create 2 NCCL communicators, one global communicator for cross-module communication (NCCLSend/Recv) and one local communicator for collective tensor communication (NCCLAllReduce) inside modules.
+We also need to handle tensor communication between modules if they have different partitions which can be naively implemented by gather and scatter.
+We use cuBLAS and cuDNN for matrix computation and convolution networks.
+To optimize memory management, we statically allocate GPU memory before the training loop and avoid time-exhausting cudaMalloc/cudaFree inside the training loop.
+\par
+ We also use some code-motion techniques to lift the cudaMalloc inside the training loop and pre-allocate the memory before the training loop to optimize memory management.
+This could significantly improve performance because cudaMalloc/Free takes a very long time.\par
+ Our generated CUDA code can efficiently represent the correct training loop of machine learning models with both Spatial Partition and Pipeline Parallelism.
+ \section{Evaluation}\label{evaluation}
+ In this section, we show the evaluation benchmarks of various forms of Pipeline parallelism by running our generated CUDA code on 4 GeForce TITAN XP GPUs or server cuda.cs.purdue.edu.
+The server is a single GPU cluster with a high-speed inter-GPU connection.
+The ideal network topologies for Pipeline parallelism is a large training system consisting of several Nodes but due to limited hardware resources, we currently could not run the evaluation under that scenario.
+The purpose of this benchmark is to validate our graph transformation and gain a deeper insight into the choice of different parallelism strategies.\par
+ We perform our test on a 12 convolution layer network for 40 iterations with different parallelism settings:
+ \begin{itemize}
+   \item Data parallelism running on 2 GPUs
+   \item Data parallelism running on 4 GPUs
+   \item Mixed parallelism running on 4 GPUs, Naive 2 module parallelism with each module running in Data parallelism on 2 GPUs
+   \item Mixed parallelism running on 4 GPUs, 2 modules in Pipelined parallelism (4 pipelines) with each module running in Data parallelism on 2 GPUs
+   \item Mixed parallelism running on 4 GPUs, 2 modules in Pipelined parallelism (8 pipelines) with each module running in Data parallelism on 2 GPUs
+ \end{itemize}
+ \begin{figure}[htbp]
+  \centering
+  \includegraphics[scale=0.5]{pipelineparallelruntime.png}
+  \caption{Data Parallelism vs Pipeline parallelism}
+  \label{fig:runtime}
+\end{figure}
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[scale=0.5]{pipelineparallelspeedup.png}
+  \caption{Data Parallelism vs Pipeline parallelism}
+  \label{fig:speedup}
+\end{figure}
+ The running time of 5 settings is shown in figure\ref{fig:runtime}, the corresponding speedup is shown in figure\ref{fig:speedup}.
+We can notice that simply dividing a model into 2 modules will not cause any speed up due to the low device usage caused by the cross-module dependency.
+The device will idle while waiting for data to arrive.
+We can notice that using Pipelined parallelism with 4 pipelines rather than naive module partition on 2 modules will cause a speedup of 1.6x.
+And we can further accelerate the training loop by increasing the pipeline number (micro-batch number) to 8 and reaching a speedup of 1.77.
+But keep in mind that increasing pipeline number will increase memory usage due to versioned tensor.\par
+ The speed-up achieved is significant and is very close to the theoretical limit.
+No matter how many pipelines we choose, there will still exist some bubble time because the different modules of the same micro-batch will still have dependencies between them.
+For 2 modules, the theoretical speedup limit is $\frac{2n}{n+1}$ ($n$ is the pipeline number), which is about the same as our evaluation result.\par
+ However, we find out that 2 modules on Pipeline parallelism with each module running at Data Parallelism on 2 GPUs (Mixed Parallelism) still perform a little worse than Data parallelism on 4 GPUs.
+This result is anticipated because we perform our evaluation on a single GPU cluster with very high-speed inter-GPU connections, which makes the frequent inter-GPU tensor communication (ALLReduce) overhead introduced by Data Parallelism negligible, but the cross-module dependencies for micro-bench will still cause some bubble time.
+As we mentioned before, the ideal network topologies for Pipeline parallelism are a large training system consisting of multiple GPU clusters connected through the network, then we can use Data parallelism inside each cluster and Pipeline parallelism across clusters.
+We currently could not perform evaluation under this ideal scenario due to limited hardware resources, but we could predict that Pipeline parallelism mixed with Data parallelism will perform much better than only using Data parallelism because Pipeline parallelism could reduce the communication overhead because inter-cluster communication only happens at module boundaries.\par
+ The evaluation shows that our framework is flexible and the user could switch to different combinations of Parallelism settings to reduce the total communication overhead and achieve optimal performance.\par
+ The evaluation shows that our framework can unify both Spatial Partition and Pipeline Parallelism, validates the correctness of our training loop transformation, and offers a deeper insight into how to choose between different parallelism training settings.
+ \section{Conclusion} \label{conclusion}
+In conclusion, we implement Pipeline parallelism for AIRCoP, make it the first machine learning framework to integrate all the parallelism strategies, and offer user flexibility to fully explore different combinations of parallel training strategies under different hardware scenarios and gain a deeper insight into parallel training.
+\bibliographystyle{acm}
+\bibliography{paper}
 \end{document}

### 2021-12-12 report/paper.tex
proof read finished

@@
 %%\documentclass[sigplan,nonacm]{acmart}
-\documentclass[sigplan]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[sigplan, nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 
@@
 \setcopyright{none}
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2021}
-\acmMonth{9}
-\acmDOI{}
+%%\acmJournal{PACMPL}
+%%\acmVolume{1}
+%%\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
+%%\acmArticle{1}
+%%\acmYear{2021}
+%%\acmMonth{9}
+%%\acmDOI{}
 %%\copyrightyear{2018}
@@
 \begin{abstract}
-  Various approaches have been developed like data parallelism, model parallelism, attribute parallelism, and pipeline parallelism to train machine learning models in parallel, however, there are no existing frameworks that can integrate these approaches.
-Most commonly used machine learning frameworks such as TensorFlow and Pytorch only support a very small subset of parallel paradigms - the data parallelism and model parallelism.
+  Various approaches have been developed like data parallelism, model parallelism, attribute parallelism, and pipeline parallelism to train machine learning models in parallel.
+However, there are no existing frameworks that can integrate these approaches.
+Most commonly used machine learning frameworks such as TensorFlow and Pytorch only support a tiny subset of parallel paradigms - the data parallelism and model parallelism.
 GPipe and PipeDream support different types of pipeline parallelism by dividing the model into several modules and running them in a pipeline fashion.
   AIRCop (Abstract IR for Comprehensive Parallelization) is a developing project that unifies all the parallelism paradigms mentioned above.
-Since all the machine learning models can be viewed as computation graphs, the parallelism paradigms (data parallelism, model parallelism, attribute parallelism, and pipeline parallelism) are just different kinds of computation graph transformation, ie transformation pass.
+Since all the machine learning models can be viewed as computation graphs, the parallelism paradigms (data parallelism, model parallelism, attribute parallelism, and pipeline parallelism) are just different kinds of computation graph transformation, i.e., transformation pass.
 AIRCop uses this idea and takes a machine learning model (basic sequential model) and parallelization annotations (describe the desired parallel behavior) as input and generates target program through different transformation passes.
@@
 \subsection{Proposal Retrospect}
-In the proposal, we mentioned that both stacked pipeline and queued pipeline will be supported by this project. However, we currently only support stacked pipeline for pipeline execution.\par
-The reason is that we spent some time adding support for multiple-module for AIRCoP because AIRCoP originally only support single module.
-This involves changing the syntax of the frontend source language to make writing multiple-module models intuitive and easy.
+In the proposal, we mentioned that this project would support both stacked pipeline and queued pipeline. However, we currently only support stacked pipeline for pipeline execution.\par
+The reason is that we spent some time adding support for multiple-module for AIRCoP because AIRCoP initially only supported single module.
+This work involves changing the syntax of the frontend to make writing multiple-module models intuitive and easy.
 After finishing support for the stacked pipeline, We also found a severe problem in AIRCoP that will cause significant performance overhead - Memory Management.
-AIRCoP originally use cudaMalloc and cudaFree inside the training loop for memory allocation.
+AIRCoP initially uses cudaMalloc and cudaFree inside the training loop for memory allocation.
 Due to the complex synchronization mechanism of cudaMalloc/Free, these operations will cause a very long time.
-After analyzing the running time with NVIDIA profiler, we found that most of the execution time is spent on cudaMalloc/Free and it could be even 10x faster by pre-allocating the memory before the training loop.
+After analyzing the running time with NVIDIA profiler, we found that most of the execution time is spent on cudaMalloc/Free, and it could be even 10x faster by pre-allocating the memory before the training loop.
 So we focus on addressing the memory management problem after finishing stacked pipeline.
@@
 Although we only support stacked pipeline, it is sufficient for multi-module pipeline parallelism.
-We can still mix spatial partition with pipeline execution to explore different parallelism settings and gain some insight into parallel training as we will see in the evaluation.\par
+We can still mix spatial partition with pipeline execution to explore different parallelism settings and gain insight into parallel training, as we will see in the evaluation.\par
 Our progress is shown as follows:
@@
 \section{Introduction} \label{introduction}
- Real-world machine learning applications are usually very large, they require a tremendous amount of data for training and can not work on a single device.
+ Real-world machine learning applications are usually tremendous.
+They require a tremendous amount of data for training and can not work on a single device.
 As a result, various parallel strategies have been proposed for running large models in parallel among multiple distributed devices to achieve optimal performance.\par
@@
 Data parallelism is adapted and developed by most modern machine learning framework (TensorFlow\cite{abadi2016tensorflow}, Pytorch\cite{paszke2019pytorch}).
-Data parallelism requires network weight synchronizations between devices at the end of each iteration, therefore Data parallelism is ideal for a computation-intensive network with few weight parameters (convolution and pooling layer) but will introduce significant communication overhead for a network with a large number of weight parameters (fully-connected layer).
+Data parallelism requires network weight synchronizations between devices at the end of each iteration.
+Therefore Data parallelism is ideal for a computation-intensive network with few weight parameters (convolution and pooling layer) but will introduce significant communication overhead for a network with a large number of weight parameters (fully-connected layer).
 Model parallelism\cite{dean2012large} is achieved by splitting the network weight tensor across devices while broadcasting the input tensor.
@@
 For example, \cite{krizhevsky2014one} uses data parallelism for the convolution layer and model parallelism for the fully-connected layer.
-All the above parallelism strategies can be categorized as Spatial Partition because they are based on the partition for data/weight/operation.
+The above parallelism strategies can be categorized as Spatial Partition because they are based on the partition for data/weight/operation.
 \par
  \subsection{Pipeline Parallelism}
- Different from Spatial Partition where input data/weight is distributed across multiple devices, there is another parallelism strategy: Pipeline Parallelism.
+ Unlike Spatial Partition, where input data/weight is distributed across multiple devices, there is another parallelism strategy: Pipeline Parallelism.
 Pipeline Parallelism divides a machine learning model into several modules with data dependency between them, divides a batch into several micro-batches, and runs them in parallel in a pipelined fashion.
 Both GPipe\cite{huang2019gpipe} and Pipedream\cite{narayanan2019pipedream} studied and developed Pipeline parallelism.
-Pipeline parallelism is especially useful when the training model is too big to fit inside a single device and there are dependencies between different layers.
-\par
- As shown in \cite{krizhevsky2014one}, combining different parallelism strategies can achieve better performance than only using one strategy.
-However, there is no unifying machine learning framework that can support all the parallelism strategies.
+Pipeline parallelism is especially useful when the training system has low-bandwidth connections because it can limit the communication overhead to only happen at module boundaries.\par
+ As shown in \cite{krizhevsky2014one}, combining different parallelism strategies can achieve better performance for different applications.
+However, no existing machine learning framework can support all the parallelism strategies.
 TensorFlow\cite{abadi2016tensorflow} and Pytorch\cite{paszke2019pytorch} only supports Data Parallelism while GPipe\cite{huang2019gpipe} and Pipedream\cite{narayanan2019pipedream} supports Pipeline parallelism.
-Users are limited to only a subset of Parallelism strategies, and can not fully utilize the distributed training system.
+Users are limited to only a subset of Parallelism strategies and can not fully utilize the distributed training system.\par
   Different distributed system topologies also would require different strategies for optimal performance.
@@
   \label{fig:networktopo}
-\end{figure} This overhead may be negligible for a GPU Cluster with high-bandwidth interconnections but will cause a significant performance drop for large distributed systems consisting of several nodes of GPU clusters with low bandwidth connection between them\ref{fig:networktopo}.
+\end{figure} This overhead may be negligible for a GPU Cluster with high-bandwidth interconnections but will cause a significant performance drop for large distributed systems consisting of several nodes of GPU clusters with low bandwidth connection between them Figure \ref{fig:networktopo}.
 Because high bandwidth inter-GPU connection techniques like NV-Link can only scale to a limited number of devices, large training servers are often deployed across multiple nodes.
-It will be ideal to divide the model into several modules and assign each module into a cluster, use Spatial partition inside a GPU cluster and adopt pipeline parallelism for modules across different clusters.\par
+It will be ideal to divide the model into several modules, assign each module to a cluster, use Spatial partition inside a GPU cluster, and adopt pipeline parallelism for modules across different clusters.\par
 \subsection{AIRCoP}
- AIRCoP \cite{aircopfeiwang} is the first framework that supports all the parallel parallelisms above thus allows users to fully explore different parallel settings.
+ AIRCoP \cite{aircopfeiwang} is the first framework that supports all the parallel parallelisms above, therefore allows users to fully explore different parallel settings.
 AIRCoP implemented spatial partition before and our work implemented Pipeline parallelism for AIRCoP, thus make AIRCoP the first unifying parallel training framework.\par
@@
   tt\in TensorType& := \mathcal{T}(shape:List[Size])\\
-\end{align*}
-\begin{align*}
+%\end{align*}
+%\begin{align*}
   \textbf{OperationType}:& \\
@@
   op \in OperationType&:= mem\_op \text{ } | \text{ } ml\_op \\
+\end{align*}
+\begin{align*}
   \textbf{Module}:& \\
@@
  AIRCoP developed a formal syntax to represent tensor computation graphs.
-The syntax mainly has 3 components: Tensor and corresponding TensorType to represent input tensor and weight, Operation, and corresponding Operation Type to represent different Machine learning operations like Conv/MatMul/Relu and their input/output Type, Module to represent different modules.
+The syntax mainly has three components: Tensor and corresponding TensorType to represent input tensor and weight, Operation, and corresponding Operation Type to represent different Machine learning operations like Conv/MatMul/Relu and their input/output Type, Module to represent different modules.
 Operations and Modules are connected through let-bindings.
-And we use annotation of the formal syntax as an approach to formally represent different parallelism strategies.
+Furthermore, we use annotation of the formal syntax to represent different parallelism strategies formally.
 We developed annotation systems for both Tensor, Operation, and Modules.
@@
 Module annotation indicates how many micro-batches the mini-batch is split into.
-Data parallelism can be achieved through annotating the input tensor indicating which dimension to split.
-Model parallelism can be achieved by annotating the weight tensor.
-Attribute parallelism can be achieved by a combination of tensor annotation and operation annotation.
-And pipeline parallelism can be simply done by annotation the Modules.
-Users only need to write the sequential forward pass of the Machine Learning model and the correct final training loop will be automatically computed through a series of transformations.
+Users only need to write the sequential forward computation of the Machine Learning model, and the correct final training loop will be automatically computed through a series of transformations.
  \section{Pipeline Parallelism} \label{pipelineparallism}
  Pipeline Parallelism is ideal for distributed training workload across several nodes with weak inter-connection.
-Simply using Spatial Partition across different nodes will cause significant communication overhead between each operation.
-By splitting the training models into several modules and assigning each module to different nodes will limit the inter-node communication to the boundary of each module, therefore significantly reducing the inter-node communication overhead.\par
+Using Spatial Partition across different nodes will cause significant communication overhead between each operation.
+Splitting the training models into several modules and assigning each module to different nodes will limit the inter-node communication to the boundary of each module, therefore significantly reducing the inter-node communication overhead.\par
  \begin{figure}[htbp]
@@
  Simply splitting modules will cause large amount of stall time (or bubble time) due to complex dependencies between the sequential modules.
-As shown in\ref{fig:multiplemodule}, if we naively spilt the model into 4 modules $m_0,m_1,m_2,m_3$.
+As shown in Figure \ref{fig:multiplemodule}, if we naively spilt the model into 4 modules $m_0,m_1,m_2,m_3$.
 There will be sequential dependencies between the forward pass $F_0, F_1, F_2, F_3$ of modules $m_0,m_1,m_2,m_3$, and reverse dependencies between backward pass $B_3, B_2, B_1, B_0$.
@@
 \end{figure}
- Our approach is similar to GPipe\cite{huang2019gpipe}, we split a batch into multiple micro-batches and run the forward pass and the backward pass of multiple micro-batches sequentially.
-Only the forward pass and backward pass of the same micro-batches have dependencies between them.
-Different modules of different micro-batches could run in parallel as shown in\ref{fig:stackedpipeline}.
-Pipeline parallelism could allow users to fully utilize distributed devices by minimizing stall time caused by cross-module dependencies.
+\subsection{Stacked Pipeline Parallelism}
+ Our approach is similar to GPipe\cite{huang2019gpipe}.
+We split a batch into multiple micro-batches and sequentially run the forward pass and the backward pass of multiple micro-batches.
+Because different micro-batches are stacked, we also call this Stacked Pipeline Parallelism.
+Only the forward and backward passes of the same micro-batches have dependencies between them.
+Different modules of different micro-batches could run in parallel as shown in Figure \ref{fig:stackedpipeline}.
+Stacked Pipeline parallelism allows users to fully utilize distributed devices by minimizing stall time caused by cross-module dependencies.
 \par
- Pipeline parallelism will insert blocking send/recv operations at the module boundaries to insure necessary execution ordering.
-And the gradient of multiple micro-batches will be accumulated after all the micro-batches finish and updated the weight synchronously for the entire mini-batch.
-Pipeline parallelism can easily scale to a large number of devices while maintaining ideal utilization of resources.\par
- Pipeline parallelism will introduce little communication overhead because it will only transfer data at the module boundaries rather than the frequent inter-operation tensor communication introduced by Spatial Partition.
-This feature of Pipeline parallelism makes it suitable not only for high-speed interconnection systems but also weakly connected systems.
+ Stacked Pipeline parallelism will insert blocking send/recv operations at the module boundaries to maintain necessary data dependencies.
+The gradient of multiple micro-batches will be accumulated and is used to update weight synchronously for the entire mini-batch.
+Stacked Pipeline parallelism can easily scale to a large number of devices while maintaining ideal utilization of resources.\par
+ Stacked Pipeline parallelism will introduce little communication overhead because it will only transfer data at the module boundaries rather than the frequent inter-operation tensor communication introduced by Spatial Partition.
+This feature of Stacked Pipeline parallelism makes it suitable for both high-speed interconnection systems and weakly connected systems.
 \par
+ Throughout the rest of this paper, the term Pipeline parallelism refers to Stacked Pipeline parallelism.
+ \subsection{Pipeline Parallelism in AIRCoP}
+ Pipeline parallelism in AIRCoP is done through module annotation systems. The annotation consists of an integer representing the number of micro-batches (or pipelines).
+ Users can split the model into several modules and annotate them for Pipeline parallelism. The resulting training loop is different from naive multiple modules.
  \begin{align*}
@@
 \end{align*}
- Pipeline parallelism in AIRCoP is done through module annotation systems.
-The annotation consists of an integer representing the number of micro-batches (or pipelines).
-Users can split the model into several modules and annotate them for Pipeline parallelism.
-The resulting training loop is slightly different from naive multiple modules.
-Assume we have $n$ pipelines, the target program will have a top-level training loop for iterating mini-batches.
-Inside each mini-batch iteration, there will be a training loop of size $n$ for iterating the forward pass of the $n$ micro-batches and a training loop of size $n$ for iterating the backward pass of the $n$ micro-batches.
+Assume we have $\epsilon$ pipelines.
+The target program will have a top-level training loop for iterating mini-batches.
+Inside each mini-batch iteration, there will be a training loop of size $\epsilon$ for iterating the forward passes of the $\epsilon$ micro-batches and a training loop of size $\epsilon$ for iterating the backward passes of the $\epsilon$ micro-batches.
 The gradient will be accumulated inside each backward pass of micro-batches.
@@
 %\end{align*}
-Transformation for pipeline parallelism has 2 stages; Collection Phase and Generation Phase.
+Transformation for pipeline parallelism has 2 stages; \textbf{Collection Phase} and \textbf{Generation Phase}.
 The Collection Phase collects the necessary information like forward/backward operation.
 The Generation Phase uses the collected information and the module annotation to generate the training loop.\par
+\paragraph{Collection Phase}
 During Collection Phase we need to collect the weight operations $E_w$, forward operations $E_f$ and backward operations $E_b$.
@@
 The corresponding variables of the operation is $xs^w$, $xs^f$, $xs^b$.
-We also need to calculate the lift variables $xs^l = Defined(E_f) \bigcap Used(E_b)$, the lifted variables $xs^l$ are the variables that are defined in the forward pass and used in the backward pass.
-Because we are generating different loops for forward pass and backward pass of micro-batches, we need to lift $xs^l$ into the outer scope and keep multiple versions of those variables for different micro-batches.\par
-The formal description of the generation phase is as follows:
+We also need to calculate the lift variables $xs^l = Defined(E_f) \bigcap Used(E_b)$, the lifted variables $xs^l$ are the variables defined in the forward pass and used in the backward pass.
+Because we are generating different loops for forward passes and backward passes of micro-batches, we need to lift $xs^l$ into the outer scope and keep multiple versions of those variables for different micro-batches.\par
+\paragraph{Generation Phase}
+The formal description of the Generation Phase is as follows:
 \begin{align*}
@@
   &\textbf{let } \phi=Assign([xs, xs^\epsilon], xs.type) \textbf{ in}\\
-  &G_f(\emptyset, xs^l)
-\end{align*}
-\begin{align*}
+  &G_f(\emptyset, xs^l)\\
   G_b&(E_b, xs::xs^f)\triangleq\\
@@
   &\textbf{let } \phi=Accumulate([t, xs^\prime], out\_type) \textbf{ in}\\
-  &G_b(E^b, \emptyset)
-\end{align*}
-\begin{align*}
+  &G_b(E^b, \emptyset)\\
   G_o&(xs::xs^w)\triangleq\\
@@
   &G_o(xs^w)\\
-  op_l&([input],out\_type) \triangleq op(input^\epsilon, out\_type)
+  \\
+  op_l&([input],out\_type) \triangleq op([input^\epsilon], out\_type)
 \end{align*}
-$op_l$ means that the operation will read from the versioned lift variable instead of the originally variable.
+$op_l$ means that the operation will read from the versioned lift variable instead of the original variable.
 $xs^\epsilon = Zero^\epsilon$ will generates an array of $\epsilon$ versions of the variable $xs$.\par
 $G_w$ generates the weight and their gradient and momentum, $G_l$ generate an array of lift nodes for different micro-batches, $G_f$ generates forward pass and write the correct value to the versioned lift variables, $G_b$ generates the backward pass and read from the versioned lift variable, $G_o$ generates the optimization pass.
-Because our source language does not allow mutation, we only need to write the versioned lift node at the end of the forward pass and the Generation phase can generate the correct training loop.
+Because our source language does not allow mutation, all the mutation operations (Assign/Accumulate/Optimize) are automatically generated in the Generation Phase.
+We only need to write the versioned lift node at the end of the forward pass, and the Generation phase can generate the correct training loop.
 \section{Implementation} \label{implementation}
- We implement AIRCoP and Pipeline Parallelism in LMS compiler framework\cite{rompf2010lightweight}, and generate CUDA\cite{sanders2010cuda} code with MPI\cite{gropp1999using}/NCCL for communication as the target executable program.\par
+We implement AIRCoP and Stacked Pipeline Parallelism in LMS compiler framework\cite{rompf2010lightweight}, and generate CUDA\cite{sanders2010cuda} code with MPI\cite{gropp1999using}/NCCL for communication as the target executable program.\par
+\subsection{Integration into LMS}
  LMS (Lightweight Modular Staging) is a library-based staging compiler that distinguishes stages solely based on Type information.
-LMS is extremely useful for developing DSL through overloading.
+LMS is beneficial for developing DSL through overloading.
 It has a typed frontend making the code generation safe.
-LMS offers lifted (next-stage) versions of normal operators and normal control-flow statements make symbolic code generation easy and convenient.
+LMS offers lifted (next-stage) versions of regular operators, and standard control-flow statements make symbolic code generation easy and convenient.
 AIRCoP is simply a tensor computation DSL embedded inside LMS framework.\par
@@
 \par
+\subsection{Training loop transformation For Pipeline Parallelism}
 We use the transformation pass of LMS to transform the IR and simulate the tensor graph transformation to generate the correct training loop.
@@
 For example, if matrix multiplication happens between matrixes of dimension $[m, n]$ and dimension $[j,k]$, then dimension $n$ and dimension $j$ can be merged.
-Annotations of tensors and operations represent which dimension to partition, so we have to update those annotations as well.
+Annotations of tensors and operations represent which dimension to partition, so we have to update those annotations too.
 \par
 The next pass DistributeTensorAIRCoP is the core pass to implement training loop transformation. This pass has 2 phases \textbf{Collection Phase} and \textbf{Generation Phase}.\par
- The \textbf{Collection Phase} collects the forward operations, the weight nodes, and the corresponding backward operation of each node.
+\paragraph{Collection Phase}
+The Collection Phase collects the forward operations, the weight nodes, and the corresponding backward operation of each node.
 The backward operation is collected in reverse order to maintain the correct semantic of differentiation.
-After Collection Phase, we have the necessary information needed to perform automatic differentiation (AD).
-To generate the correct training loop, we also need to collect lifted tensors.
+Because the backward pass should be generated after the forward pass, we collect backward operation in the form of closures.
+We implement the Collection Phase by defining a function aircopCollect for each IR Node to collect the corresponding information.
+After the Collection Phase, we have the necessary information to perform automatic differentiation (AD).
+We also need to collect lifted tensors to generate the correct training loop.
 As we mentioned before, the forward pass and backward pass of micro-batches are iterated in different loops to achieve pipeline parallelism.
-Therefore, the tensors that are allocated in the forward pass and used in the backward pass will cause an error because forward and backward passes are now in different scopes.
-These tensors are called lifted tensors, we have to lift them into the outer scope and maintain multiple versions of them as an array of tensors.
-The forward and backward pass of micro-batches will use different versions of the lifted tensors.
+Therefore, the tensors allocated in the forward pass and used in the backward pass will cause an error because forward and backward passes are now in different scopes.
+These tensors are called lifted tensors.
+We have to lift them into the outer scope and maintain multiple versions of them as an array of tensors.
+Micro-batches' forward and backward passes will use different versions of the lifted tensors.
 After collecting the lifted tensors, the Collection Phase is finished.\par
- The \textbf{Generation Phase} generates the correct training loop based on the collected information.
+\paragraph{Generation Phase}
+The Generation Phase generates the correct training loop based on the collected information.
 It will add paired send/recv operations at the boundaries of different modules to maintain the correct data dependencies.
 Inside each module, it will first generate the top-level training loop for mini-batches.
-Then it will generate the forward training loop and backward training loop for micro-batches based on Module Annotations and use the collected backward operations to realize automatic differentiation.
+Then it will generate the forward training loop and backward training loop for micro-batches based on Module Annotations and use the collected backward operations to realize automatic differentiation as we mentioned in Section \ref{formalpipe}.
 Inside both the forward and backward pass, the read and write to the lifted tensors will be transformed into read/write to the versioned tensor array.
-The weight gradient will be accumulated in the backward pass and optimized in the end for the entire mini-batch.\par
+The weight gradient will be accumulated in the backward pass and optimized at the end of the entire mini-batch.\par
+\subsection{Spatial Partition Pass}
 The next pass will perform the Spatial Partition transformation.
-This part is of past work so we will not dig further into this pass.
+This part is of past work, so we will not dig further into this pass.
 The main purpose of this pass is to partition the tensor and operations based on annotations and add the corresponding inter operation tensor communication for synchronization (AllReduce for MatrixMultiplication or Convolution).\par
+\subsection{MPI/NCCL Pass}
 The final pass is to generate the corresponding MPI/NCCL instructions for inter-device communications.
 NCCL (NVIDIA Collective Communications Library) is the communication library introduced by NVIDIA for efficient collective and point-to-point communication across multiple GPU and multiple Nodes.
-Our target program is CUDA code with MPI/NCCL for communication, we use MPI to set up the environment and the NCCL communicator.
+Our target program is CUDA code with MPI/NCCL for communication.
+We use MPI to set up the environment and the NCCL communicator.
 We will create 2 NCCL communicators, one global communicator for cross-module communication (NCCLSend/Recv) and one local communicator for collective tensor communication (NCCLAllReduce) inside modules.
-We also need to handle tensor communication between modules if they have different partitions which can be naively implemented by gather and scatter.
-We use cuBLAS and cuDNN for matrix computation and convolution networks.
-To optimize memory management, we statically allocate GPU memory before the training loop and avoid time-exhausting cudaMalloc/cudaFree inside the training loop.
-\par
- We also use some code-motion techniques to lift the cudaMalloc inside the training loop and pre-allocate the memory before the training loop to optimize memory management.
-This could significantly improve performance because cudaMalloc/Free takes a very long time.\par
+We also need to handle tensor communication between modules if they have different Spatial partitions, which can be naively implemented by gather and scatter.
+We use cuBLAS and cuDNN for matrix computation and convolution networks.\par
+We statically allocate GPU memory before the training loop and avoid time-exhausting cudaMalloc/cudaFree inside the training loop to optimize memory management.
+We currently use code-motion techniques to lift the cudaMalloc inside the training loop and pre-allocate the memory before the training loop to optimize memory management.
+This approach could significantly improve performance because cudaMalloc/Free often takes a very long time.\par
 Our generated CUDA code can efficiently represent the correct training loop of machine learning models with both Spatial Partition and Pipeline Parallelism.
  \section{Evaluation}\label{evaluation}
- In this section, we show the evaluation benchmarks of various forms of Pipeline parallelism by running our generated CUDA code on 4 GeForce TITAN XP GPUs or server cuda.cs.purdue.edu.
+ This section shows the evaluation benchmarks of various forms of Pipeline parallelism by running our generated CUDA code on 4 GeForce TITAN XP GPUs or server cuda.cs.purdue.edu.
 The server is a single GPU cluster with a high-speed inter-GPU connection.
-The ideal network topologies for Pipeline parallelism is a large training system consisting of several Nodes but due to limited hardware resources, we currently could not run the evaluation under that scenario.
+The ideal network topology for Pipeline parallelism is an extensive training system consisting of several Nodes, but we currently could not run the evaluation under that scenario due to limited hardware resources and time.
 The purpose of this benchmark is to validate our graph transformation and gain a deeper insight into the choice of different parallelism strategies.\par
- We perform our test on a 12 convolution layer network for 40 iterations with different parallelism settings:
+ We perform our test on a 12 layer convolution network for 40 iterations with different parallelism settings:
  \begin{itemize}
@@
    \item Data parallelism running on 4 GPUs
-   \item Mixed parallelism running on 4 GPUs, Naive 2 module parallelism with each module running in Data parallelism on 2 GPUs
-   \item Mixed parallelism running on 4 GPUs, 2 modules in Pipelined parallelism (4 pipelines) with each module running in Data parallelism on 2 GPUs
-   \item Mixed parallelism running on 4 GPUs, 2 modules in Pipelined parallelism (8 pipelines) with each module running in Data parallelism on 2 GPUs
+   \item Mixed parallelism running on 4 GPUs --- Naively partitioned into 2 modules with each module running in Data parallelism on 2 GPUs
+   \item Mixed parallelism running on 4 GPUs --- 2 modules in Stacked Pipeline parallelism (4 pipelines) with each module running in Data parallelism on 2 GPUs
+   \item Mixed parallelism running on 4 GPUs --- 2 modules in Stacked Pipeline parallelism (8 pipelines) with each module running in Data parallelism on 2 GPUs
  \end{itemize}
@@
  The running time of 5 settings is shown in figure\ref{fig:runtime}, the corresponding speedup is shown in figure\ref{fig:speedup}.
-We can notice that simply dividing a model into 2 modules will not cause any speed up due to the low device usage caused by the cross-module dependency.
+We can notice that simply dividing a model into 2 modules will not cause any speed up due to the low device usage caused by the cross-module dependencies.
 The device will idle while waiting for data to arrive.
-We can notice that using Pipelined parallelism with 4 pipelines rather than naive module partition on 2 modules will cause a speedup of 1.6x.
-And we can further accelerate the training loop by increasing the pipeline number (micro-batch number) to 8 and reaching a speedup of 1.77.
-But keep in mind that increasing pipeline number will increase memory usage due to versioned tensor.\par
- The speed-up achieved is significant and is very close to the theoretical limit.
-No matter how many pipelines we choose, there will still exist some bubble time because the different modules of the same micro-batch will still have dependencies between them.
+We can notice that using Pipeline parallelism with 4 pipelines rather than naive module partition on 2 modules will cause a speedup of 1.6x.
+Moreover, we can further accelerate the training loop by increasing the pipeline number (micro-batch number) to 8 and reaching a speedup of 1.77x.
+However, keep in mind that increasing pipeline number will increase memory usage due to versioned tensor.\par
+ The speedup achieved is significant and is very close to the theoretical limit.
+No matter how many pipelines we choose, some bubble time will still exist because the different modules of the same micro-batch will still have dependencies between them.
 For 2 modules, the theoretical speedup limit is $\frac{2n}{n+1}$ ($n$ is the pipeline number), which is about the same as our evaluation result.\par
- However, we find out that 2 modules on Pipeline parallelism with each module running at Data Parallelism on 2 GPUs (Mixed Parallelism) still perform a little worse than Data parallelism on 4 GPUs.
-This result is anticipated because we perform our evaluation on a single GPU cluster with very high-speed inter-GPU connections, which makes the frequent inter-GPU tensor communication (ALLReduce) overhead introduced by Data Parallelism negligible, but the cross-module dependencies for micro-bench will still cause some bubble time.
-As we mentioned before, the ideal network topologies for Pipeline parallelism are a large training system consisting of multiple GPU clusters connected through the network, then we can use Data parallelism inside each cluster and Pipeline parallelism across clusters.
-We currently could not perform evaluation under this ideal scenario due to limited hardware resources, but we could predict that Pipeline parallelism mixed with Data parallelism will perform much better than only using Data parallelism because Pipeline parallelism could reduce the communication overhead because inter-cluster communication only happens at module boundaries.\par
- The evaluation shows that our framework is flexible and the user could switch to different combinations of Parallelism settings to reduce the total communication overhead and achieve optimal performance.\par
- The evaluation shows that our framework can unify both Spatial Partition and Pipeline Parallelism, validates the correctness of our training loop transformation, and offers a deeper insight into how to choose between different parallelism training settings.
+ However, we find out that 2 modules on Stacked Pipeline parallelism with each module running at Data Parallelism on 2 GPUs (Mixed Parallelism) still perform a little worse than Data parallelism on 4 GPUs.
+This result is anticipated because we perform our evaluation on a single GPU cluster with very high-speed inter-GPU connections, which makes the frequent tensor communication (ALLReduce) overhead introduced by Data Parallelism negligible.
+However, the cross-module dependencies for micro-batch will still cause some bubble time.
+As we mentioned before, the ideal network topologies for Pipeline parallelism are an extensive training system consisting of multiple GPU clusters connected through the network.
+Then we can use Data parallelism inside each cluster and Pipeline parallelism across clusters.
+We currently could not perform evaluation under this ideal scenario due to limited time and hardware resources, but we could predict that Pipeline parallelism mixed with Data parallelism will perform much better than only using Data parallelism in this scenario because Pipeline parallelism could reduce the communication overhead because inter-cluster communication only happens at module boundaries.\par
+ The evaluation shows that our framework is flexible, and the user could switch to different combinations of Parallelism settings to reduce the total communication overhead and achieve optimal performance.
+It also demonstrates that our framework can unify both Spatial Partition and Pipeline Parallelism, validates the correctness of our training loop transformation, and offers a deeper insight into how to choose between different parallelism training settings.
  \section{Conclusion} \label{conclusion}
-In conclusion, we implement Pipeline parallelism for AIRCoP, make it the first machine learning framework to integrate all the parallelism strategies, and offer user flexibility to fully explore different combinations of parallel training strategies under different hardware scenarios and gain a deeper insight into parallel training.
+In conclusion, we implement Pipeline parallelism for AIRCoP, make it a machine learning framework that can integrate different parallelism strategies, and offer user flexibility to fully explore different combinations of parallel training strategies under different hardware scenarios and gain a deeper insight into parallel training.
 \bibliographystyle{acm}

### 2021-12-12 report/paper.tex
final version ready

@@
 In section \ref{implementation}, we present the implementation detail of our work.\par
-In section \ref{evaluation}, we present a parallel training evaluation with different parallelism settings to illustrate the flexibility of our framework.
+In section \ref{evaluation}, we present a parallel training evaluation with different parallelism settings to illustrate the flexibility of our framework.\par
+In section \ref{conclusion}, we summarize our progress and draw some conclusions.
 \section{Introduction} \label{introduction}
@@
   tt\in TensorType& := \mathcal{T}(shape:List[Size])\\
-%\end{align*}
-%\begin{align*}
   \textbf{OperationType}:& \\
@@
   \centering
-  \includegraphics[scale=0.3]{NormalPipeline}
+  \includegraphics[scale=0.4]{NormalPipeline}
   \caption{Naive multiple modules}
@@
   \centering
-  \includegraphics[scale=0.4]{stackedpipeline}
+  \includegraphics[scale=0.45]{stackedpipeline}
   \caption{Pipeline Parallelism}
@@
 The corresponding variables of the operation is $xs^w$, $xs^f$, $xs^b$.
-We also need to calculate the lift variables $xs^l = Defined(E_f) \bigcap Used(E_b)$, the lifted variables $xs^l$ are the variables defined in the forward pass and used in the backward pass.
+We also need to calculate the lift variables $$xs^l = Defined(E_f) \bigcap Used(E_b)$$, the lifted variables $xs^l$ are the variables defined in the forward pass and used in the backward pass.
 Because we are generating different loops for forward passes and backward passes of micro-batches, we need to lift $xs^l$ into the outer scope and keep multiple versions of those variables for different micro-batches.\par

### 2021-12-12 report/paper.tex
Second proof read and Minor Grammar fix

@@
   \label{fig:networktopo}
-\end{figure} This overhead may be negligible for a GPU Cluster with high-bandwidth interconnections but will cause a significant performance drop for large distributed systems consisting of several nodes of GPU clusters with low bandwidth connection between them Figure \ref{fig:networktopo}.
+\end{figure} This overhead may be negligible for a GPU Cluster with high-bandwidth interconnections but will cause a significant performance drop for large distributed systems consisting of several nodes of GPU clusters with low bandwidth connection between them as shown in Figure \ref{fig:networktopo}.
 Because high bandwidth inter-GPU connection techniques like NV-Link can only scale to a limited number of devices, large training servers are often deployed across multiple nodes.
-It will be ideal to divide the model into several modules, assign each module to a cluster, use Spatial partition inside a GPU cluster, and adopt pipeline parallelism for modules across different clusters.\par
+It will be ideal to divide the model into several modules, assign each module to a cluster, use Spatial partition inside a GPU cluster, and adopt pipeline parallelism between modules.\par
 \subsection{AIRCoP}
- AIRCoP \cite{aircopfeiwang} is the first framework that supports all the parallel parallelisms above, therefore allows users to fully explore different parallel settings.
-AIRCoP implemented spatial partition before and our work implemented Pipeline parallelism for AIRCoP, thus make AIRCoP the first unifying parallel training framework.\par
+ AIRCoP \cite{aircopfeiwang} is a machine learning framework that supports all the parallel parallelisms above, therefore allows users to fully explore different parallel settings.
+AIRCoP implemented spatial partition before and our work implemented Pipeline parallelism for AIRCoP, thus make AIRCoP a unifying parallel training framework.\par
  %\begin{figure}[htbp]
@@
 The syntax mainly has three components: Tensor and corresponding TensorType to represent input tensor and weight, Operation, and corresponding Operation Type to represent different Machine learning operations like Conv/MatMul/Relu and their input/output Type, Module to represent different modules.
-Operations and Modules are connected through let-bindings.
+Operations are connected through let-bindings.
 Furthermore, we use annotation of the formal syntax to represent different parallelism strategies formally.
@@
 There will be sequential dependencies between the forward pass $F_0, F_1, F_2, F_3$ of modules $m_0,m_1,m_2,m_3$, and reverse dependencies between backward pass $B_3, B_2, B_1, B_0$.
-The dependencies between the forward pass and backward pass will force strict ordering between modules, introduce synchronizations between module boundaries, and cause significant idle time on 4 devices.
+The dependencies between the forward pass and backward pass will force strict ordering between modules, introduce synchronizations at module boundaries, and cause significant idle time on 4 devices.
 \par
@@
 It has a typed frontend making the code generation safe.
-LMS offers lifted (next-stage) versions of regular operators, and standard control-flow statements make symbolic code generation easy and convenient.
+LMS offers lifted (next-stage) versions of regular operators and standard control-flow statements, making symbolic code generation easy and convenient.
 AIRCoP is simply a tensor computation DSL embedded inside LMS framework.\par
@@
 Because the backward pass should be generated after the forward pass, we collect backward operation in the form of closures.
+The Collection Phase will also collect paired send/recv operations at the boundaries of different modules to maintain the correct data dependencies.
 We implement the Collection Phase by defining a function aircopCollect for each IR Node to collect the corresponding information.
@@
 The Generation Phase generates the correct training loop based on the collected information.
-It will add paired send/recv operations at the boundaries of different modules to maintain the correct data dependencies.
 Inside each module, it will first generate the top-level training loop for mini-batches.
@@
  \section{Evaluation}\label{evaluation}
- This section shows the evaluation benchmarks of various forms of Pipeline parallelism by running our generated CUDA code on 4 GeForce TITAN XP GPUs or server cuda.cs.purdue.edu.
+ This section shows the evaluation benchmarks of various forms of parallel training by running our generated CUDA code on 4 GeForce TITAN XP GPUs on server cuda.cs.purdue.edu.
 The server is a single GPU cluster with a high-speed inter-GPU connection.
@@
 \end{figure}
- The running time of 5 settings is shown in figure\ref{fig:runtime}, the corresponding speedup is shown in figure\ref{fig:speedup}.
+ The running time of 5 settings is shown in Figure \ref{fig:runtime}, the corresponding speedup is shown in Figure \ref{fig:speedup}.
 We can notice that simply dividing a model into 2 modules will not cause any speed up due to the low device usage caused by the cross-module dependencies.
@@
 As we mentioned before, the ideal network topologies for Pipeline parallelism are an extensive training system consisting of multiple GPU clusters connected through the network.
-Then we can use Data parallelism inside each cluster and Pipeline parallelism across clusters.
+Then we can use Data parallelism inside each cluster and Pipeline parallelism between clusters.
 We currently could not perform evaluation under this ideal scenario due to limited time and hardware resources, but we could predict that Pipeline parallelism mixed with Data parallelism will perform much better than only using Data parallelism in this scenario because Pipeline parallelism could reduce the communication overhead because inter-cluster communication only happens at module boundaries.\par
  The evaluation shows that our framework is flexible, and the user could switch to different combinations of Parallelism settings to reduce the total communication overhead and achieve optimal performance.
-It also demonstrates that our framework can unify both Spatial Partition and Pipeline Parallelism, validates the correctness of our training loop transformation, and offers a deeper insight into how to choose between different parallelism training settings.
+It also demonstrates that our framework can unify both Spatial Partition and Pipeline Parallelism, validates the correctness of our training loop transformation, and offers a deeper insight into how to choose between different parallel training settings.
  \section{Conclusion} \label{conclusion}

### 2021-12-18 report/paper.tex
Spelling/grammar pass finished

@@
 \begin{abstract}
-  Various approaches have been developed like data parallelism, model parallelism, attribute parallelism, and pipeline parallelism to train machine learning models in parallel.
-However, there are no existing frameworks that can integrate these approaches.
-Most commonly used machine learning frameworks such as TensorFlow and Pytorch only support a tiny subset of parallel paradigms - the data parallelism and model parallelism.
+  Various approaches like data parallelism, model parallelism, attribute parallelism, and pipeline parallelism have been developed to train machine learning models in parallel.
+However, no existing frameworks have integrated these approaches successfully.
+Most generally used machine learning frameworks such as TensorFlow and Pytorch only support a tiny subset of parallel paradigms, i.e., data parallelism and model parallelism.
 GPipe and PipeDream support different types of pipeline parallelism by dividing the model into several modules and running them in a pipeline fashion.
   AIRCop (Abstract IR for Comprehensive Parallelization) is a developing project that unifies all the parallelism paradigms mentioned above.
-Since all the machine learning models can be viewed as computation graphs, the parallelism paradigms (data parallelism, model parallelism, attribute parallelism, and pipeline parallelism) are just different kinds of computation graph transformation, i.e., transformation pass.
-AIRCop uses this idea and takes a machine learning model (basic sequential model) and parallelization annotations (describe the desired parallel behavior) as input and generates target program through different transformation passes.
+Since all the machine learning models can be viewed as computation graphs, parallelism paradigms are just different kinds of computation graph transformation, i.e., transformation pass.
+AIRCop uses this idea, takes a machine learning model and parallelization annotations that describe the desired parallel behavior as input, and generates target program through different transformation passes.
 
@@
 \subsection{Proposal Retrospect}
-In the proposal, we mentioned that this project would support both stacked pipeline and queued pipeline. However, we currently only support stacked pipeline for pipeline execution.\par
-The reason is that we spent some time adding support for multiple-module for AIRCoP because AIRCoP initially only supported single module.
+The proposal mentions that this project will support both stacked pipeline and queued pipeline.
+However, we currently only support stacked pipeline for pipeline execution.
+The reason is that we spend some time adding support for multiple-module for AIRCoP because AIRCoP initially only support single module.
 This work involves changing the syntax of the frontend to make writing multiple-module models intuitive and easy.
-After finishing support for the stacked pipeline, We also found a severe problem in AIRCoP that will cause significant performance overhead - Memory Management.
+After finishing support for the stacked pipeline, we also find that the Memory Management problem in AIRCoP will cause a significant performance overhead.
 AIRCoP initially uses cudaMalloc and cudaFree inside the training loop for memory allocation.
-Due to the complex synchronization mechanism of cudaMalloc/Free, these operations will cause a very long time.
-After analyzing the running time with NVIDIA profiler, we found that most of the execution time is spent on cudaMalloc/Free, and it could be even 10x faster by pre-allocating the memory before the training loop.
-So we focus on addressing the memory management problem after finishing stacked pipeline.
-We currently use some code-motion techniques that will lift all the cudaMalloc out of the training loop for optimization.\par
-Because queued pipeline parallelism uses asynchronous weight update and will use lagged network weight, we choose to address the memory management problem first due to limited time.\par
-Because we only currently implement stacked pipeline, all the pipeline parallelism used in this paper refers to stacked pipeline parallelism.
+Due to the complex synchronization mechanism of cudaMalloc/Free, these operations will occupy a very long time.
+After analyzing the running time with NVIDIA profiler, we find that most of the execution time is spent on cudaMalloc/Free.
+It is even 10x faster by pre-allocating the memory before the training loop, so we focus on addressing the memory management problem after finishing stacked pipeline.
+We use code-motion techniques to lift all the cudaMalloc out of the training loop for optimization.\par
+ Queued pipeline parallelism performs asynchronous weight updates and uses lagged network weight, so we choose to address the memory management problem first due to limited time.\par
+All the pipeline parallelism used in this paper refers to stacked pipeline parallelism because we currently only implement stacked pipeline.
 \subsection{Progress}
 Although we only support stacked pipeline, it is sufficient for multi-module pipeline parallelism.
-We can still mix spatial partition with pipeline execution to explore different parallelism settings and gain insight into parallel training, as we will see in the evaluation.\par
+We can still mix spatial partition with pipeline execution to explore different parallelism settings and gain insight into parallel training, which we will present in the evaluation.\par
 Our progress is shown as follows:
@@
   \item We formalize the transformation procedure of stacked pipeline as training loop transformation on the formal language of AIRCoP.
-  \item We implemented stacked pipeline for AIRCoP.
+  \item We implement stacked pipeline for AIRCoP.
   \item We optimize the memory management for AIRCoP.
-  \item We run evaluation benchmarks with different parallelism settings(Spatial Partition, Spatial Partition mixed with Pipeline Execution) to validate our implementation and gain some insight into parallel training.
-\end{itemize} \par
+  \item We run evaluation benchmarks with different parallelism settings (Spatial Partition, Spatial Partition mixed with Pipeline Execution) to validate our implementation and gain some insight into parallel training.
+\end{itemize}
+The paper is organized as follows: \par \par
 In section \ref{introduction}, we introduce parallel training, different existing parallelism strategies, and AIRCoP.\par
 In section \ref{pipelineparallism}, we further study stacked pipeline - the pipeline parallelism we support in AIRCoP.\par
-In section \ref{formalpipe}, we show the formal description of the training loop transformation for pipeline parallelism.\par
+In section \ref{formalpipe}, we demonstrate the formal description of the training loop transformation for pipeline parallelism.\par
 In section \ref{implementation}, we present the implementation detail of our work.\par
-In section \ref{evaluation}, we present a parallel training evaluation with different parallelism settings to illustrate the flexibility of our framework.\par
+In section \ref{evaluation}, we offer a parallel training evaluation with different parallelism settings to illustrate the flexibility of our framework.\par
 In section \ref{conclusion}, we summarize our progress and draw some conclusions.
 \section{Introduction} \label{introduction}
- Real-world machine learning applications are usually tremendous.
-They require a tremendous amount of data for training and can not work on a single device.
-As a result, various parallel strategies have been proposed for running large models in parallel among multiple distributed devices to achieve optimal performance.\par
- \subsection{Spatial Partition}There are different kinds of parallelism strategies, the most common one is Data parallelism\cite{krizhevsky2012imagenet}, where we split input across multiple devices and maintain the same network weight.
-Data parallelism is adapted and developed by most modern machine learning framework (TensorFlow\cite{abadi2016tensorflow}, Pytorch\cite{paszke2019pytorch}).
-Data parallelism requires network weight synchronizations between devices at the end of each iteration.
-Therefore Data parallelism is ideal for a computation-intensive network with few weight parameters (convolution and pooling layer) but will introduce significant communication overhead for a network with a large number of weight parameters (fully-connected layer).
-Model parallelism\cite{dean2012large} is achieved by splitting the network weight tensor across devices while broadcasting the input tensor.
-Model parallelism reduces the communication overhead of Data Parallelism but is rarely used because there are often dependencies between network layers.
+Real-world machine learning applications are usually enormous.
+They require excessive data for training and cannot work on a single device.
+Therefore, various parallel strategies have been proposed for running large models in parallel among multiple distributed devices to achieve optimal performance.\par
+ \subsection{Spatial Partition}
+ The most common strategy for parallel training is Spatial Partition, where the data/weight/operation is partitioned across multiple devices.
+Among different Spatial Partition parallelism strategies, Data Parallelism\cite{krizhevsky2012imagenet}, where we split input across multiple devices and maintain the same network weight, is the simplest one.
+Data Parallelism is adapted and developed by most modern machine learning frameworks, e.g., TensorFlow\cite{abadi2016tensorflow} and Pytorch\cite{paszke2019pytorch}.
+Data Parallelism requires network weight synchronizations amid devices at the end of each iteration.
+Therefore, Data Parallelism is ideal for a computation-intensive network with few weight parameters, e.g., convolution and pooling layer, but will introduce significant communication overhead for a network with heavy weight parameters like fully-connected layers.\par
+  Model parallelism\cite{dean2012large} is another Spatial Partition based parallelism strategy and is achieved by splitting the network weight tensor across devices while broadcasting the input tensor.
+Model parallelism reduces the communication overhead of Data Parallelism but is rarely used because dependencies often exist between network layers.
 Attribute parallelism\cite{jia2019beyond} can partition the attribute dimension in both input and weight parameters and require collective communication like AllReduce between operations.
-An experienced developer can combine different Parallelism strategies to achieve optimal performance.
-For example, \cite{krizhevsky2014one} uses data parallelism for the convolution layer and model parallelism for the fully-connected layer.
-The above parallelism strategies can be categorized as Spatial Partition because they are based on the partition for data/weight/operation.
-\par
+An experienced developer can combine different Parallelism strategies to achieve optimal performance, e.g., a previous work\cite{krizhevsky2014one} uses data parallelism for the convolution layer and model parallelism for the fully-connected layer.\par
  \subsection{Pipeline Parallelism}
- Unlike Spatial Partition, where input data/weight is distributed across multiple devices, there is another parallelism strategy: Pipeline Parallelism.
-Pipeline Parallelism divides a machine learning model into several modules with data dependency between them, divides a batch into several micro-batches, and runs them in parallel in a pipelined fashion.
-Both GPipe\cite{huang2019gpipe} and Pipedream\cite{narayanan2019pipedream} studied and developed Pipeline parallelism.
-Pipeline parallelism is especially useful when the training system has low-bandwidth connections because it can limit the communication overhead to only happen at module boundaries.\par
+ Unlike Spatial Partition, where input data/weight is distributed across multiple devices, Pipeline Parallelism splits a machine learning model into several modules with data dependency between them while dividing a batch into several micro-batches.
+It then runs different modules in parallel in a pipelined fashion.
+Both GPipe\cite{huang2019gpipe} and Pipedream\cite{narayanan2019pipedream} study and develop Pipeline Parallelism.
+Pipeline Parallelism is especially useful when the training system has low-bandwidth connections because it can limit the communication overhead to only happen at module boundaries.\par
  As shown in \cite{krizhevsky2014one}, combining different parallelism strategies can achieve better performance for different applications.
 However, no existing machine learning framework can support all the parallelism strategies.
-TensorFlow\cite{abadi2016tensorflow} and Pytorch\cite{paszke2019pytorch} only supports Data Parallelism while GPipe\cite{huang2019gpipe} and Pipedream\cite{narayanan2019pipedream} supports Pipeline parallelism.
-Users are limited to only a subset of Parallelism strategies and can not fully utilize the distributed training system.\par
+TensorFlow\cite{abadi2016tensorflow} and Pytorch\cite{paszke2019pytorch} only support Data Parallelism while GPipe\cite{huang2019gpipe} and Pipedream\cite{narayanan2019pipedream} only implement Pipeline Parallelism.
+Users are limited to a subset of Parallelism strategies and cannot fully utilize the distributed training system.\par
   Different distributed system topologies also would require different strategies for optimal performance.
@@
   \label{fig:networktopo}
-\end{figure} This overhead may be negligible for a GPU Cluster with high-bandwidth interconnections but will cause a significant performance drop for large distributed systems consisting of several nodes of GPU clusters with low bandwidth connection between them as shown in Figure \ref{fig:networktopo}.
-Because high bandwidth inter-GPU connection techniques like NV-Link can only scale to a limited number of devices, large training servers are often deployed across multiple nodes.
-It will be ideal to divide the model into several modules, assign each module to a cluster, use Spatial partition inside a GPU cluster, and adopt pipeline parallelism between modules.\par
+\end{figure} This overhead may be negligible for a GPU Cluster with high-bandwidth interconnections but will cause a significant performance drop for large distributed systems consisting of several GPU clusters with low bandwidth connections among them as shown in Figure \ref{fig:networktopo}.
+Large training servers are often deployed across multiple nodes because high bandwidth inter-GPU connection techniques like NV-Link can only scale to a limited number of devices.
+It will be ideal if we can divide the model into several modules, assign each module to a cluster, use Spatial Partition inside a GPU cluster, and adopt Pipeline Parallelism among modules.\par
 \subsection{AIRCoP}
- AIRCoP \cite{aircopfeiwang} is a machine learning framework that supports all the parallel parallelisms above, therefore allows users to fully explore different parallel settings.
-AIRCoP implemented spatial partition before and our work implemented Pipeline parallelism for AIRCoP, thus make AIRCoP a unifying parallel training framework.\par
+AIRCoP \cite{aircopfeiwang} is a machine learning framework that supports all the parallel parallelisms above, thus allowing users to explore different parallel settings.
+AIRCoP has previously implemented Spatial Partition, and our work implements Pipeline Parallelism for AIRCoP, thus making AIRCoP a unifying parallel training framework.\par
  %\begin{figure}[htbp]
@@
 \end{align*}
- AIRCoP developed a formal syntax to represent tensor computation graphs.
-The syntax mainly has three components: Tensor and corresponding TensorType to represent input tensor and weight, Operation, and corresponding Operation Type to represent different Machine learning operations like Conv/MatMul/Relu and their input/output Type, Module to represent different modules.
+AIRCoP develops a formal syntax to represent tensor computation graphs.
+The syntax mainly has three components: 1.
+Tensor and corresponding TensorType to represent input tensor and weight.
+2.
+Operation and corresponding Operation Type to represent different Machine learning operations like Conv/MatMul/Relu and their input/output Type.
+3.
+Module to represent different modules.\par
 Operations are connected through let-bindings.
-Furthermore, we use annotation of the formal syntax to represent different parallelism strategies formally.
-We developed annotation systems for both Tensor, Operation, and Modules.
-Tensor and Operation annotation represents Spatial Partition by indicating the dimension to partition.
-Module annotation indicates how many micro-batches the mini-batch is split into.
+Furthermore, we use annotation of the formal syntax to represent different parallelism strategies.
+We develop annotation systems for both Tensor, Operation, and Module.
+Tensor and Operation annotation represents Spatial Partition by indicating the dimension to partition, and Module annotation represents the number of micro-batches.
 Users only need to write the sequential forward computation of the Machine Learning model, and the correct final training loop will be automatically computed through a series of transformations.
  \section{Pipeline Parallelism} \label{pipelineparallism}
- Pipeline Parallelism is ideal for distributed training workload across several nodes with weak inter-connection.
-Using Spatial Partition across different nodes will cause significant communication overhead between each operation.
-Splitting the training models into several modules and assigning each module to different nodes will limit the inter-node communication to the boundary of each module, therefore significantly reducing the inter-node communication overhead.\par
+ Pipeline Parallelism is ideal for distributed training among several nodes with weak inter-connections.
+Directly using Spatial Partition among different nodes will cause significant communication overhead between each operation.
+Splitting the training models into several modules and assigning each module to different nodes will limit the inter-node communication to the boundary of each module, thus reducing the inter-node communication overhead significantly.\par
  \begin{figure}[htbp]
@@
 \end{figure}
- Simply splitting modules will cause large amount of stall time (or bubble time) due to complex dependencies between the sequential modules.
-As shown in Figure \ref{fig:multiplemodule}, if we naively spilt the model into 4 modules $m_0,m_1,m_2,m_3$.
-There will be sequential dependencies between the forward pass $F_0, F_1, F_2, F_3$ of modules $m_0,m_1,m_2,m_3$, and reverse dependencies between backward pass $B_3, B_2, B_1, B_0$.
+Due to complex dependencies between the sequential modules, simply splitting the model will cause excessive stall time, a.k.a.
+bubble time.
+As shown in Figure \ref{fig:multiplemodule}, if we split the model into 4 modules $m_0,m_1,m_2,m_3$, there will be sequential dependencies between the forward pass $F_0, F_1, F_2, F_3$ of modules $m_0,m_1,m_2,m_3$, and reverse dependencies between backward pass $B_3, B_2, B_1, B_0$.
 The dependencies between the forward pass and backward pass will force strict ordering between modules, introduce synchronizations at module boundaries, and cause significant idle time on 4 devices.
@@
 Our approach is similar to GPipe\cite{huang2019gpipe}.
-We split a batch into multiple micro-batches and sequentially run the forward pass and the backward pass of multiple micro-batches.
-Because different micro-batches are stacked, we also call this Stacked Pipeline Parallelism.
-Only the forward and backward passes of the same micro-batches have dependencies between them.
+We split a batch into multiple micro-batches and sequentially run the forward and backward pass of multiple micro-batches.
+We call this Stacked Pipeline Parallelism because different micro-batches are stacked.
+Dependencies now only exist between the forward and backward pass of the same micro-batch.
 Different modules of different micro-batches could run in parallel as shown in Figure \ref{fig:stackedpipeline}.
-Stacked Pipeline parallelism allows users to fully utilize distributed devices by minimizing stall time caused by cross-module dependencies.
+Stacked Pipeline Parallelism allows users to fully utilize distributed devices by minimizing stall time caused by cross-module dependencies.
 \par
- Stacked Pipeline parallelism will insert blocking send/recv operations at the module boundaries to maintain necessary data dependencies.
-The gradient of multiple micro-batches will be accumulated and is used to update weight synchronously for the entire mini-batch.
-Stacked Pipeline parallelism can easily scale to a large number of devices while maintaining ideal utilization of resources.\par
- Stacked Pipeline parallelism will introduce little communication overhead because it will only transfer data at the module boundaries rather than the frequent inter-operation tensor communication introduced by Spatial Partition.
-This feature of Stacked Pipeline parallelism makes it suitable for both high-speed interconnection systems and weakly connected systems.
+Stacked Pipeline Parallelism will insert blocking send/recv operations at the module boundaries to maintain necessary data dependencies.
+The gradient of multiple micro-batches will be accumulated and used to update weight synchronously for the entire mini-batch.
+Stacked Pipeline Parallelism can easily scale to a large number of devices while maintaining ideal utilization of resources.\par
+Stacked Pipeline Parallelism will introduce little communication overhead because it will only transfer data at the module boundaries.
+This feature of Stacked Pipeline Parallelism makes it suitable both for systems with high-speed interconnections and weakly connected systems.
 \par
- Throughout the rest of this paper, the term Pipeline parallelism refers to Stacked Pipeline parallelism.
+Throughout the rest of this paper, the term Pipeline Parallelism refers to Stacked Pipeline Parallelism.
  \subsection{Pipeline Parallelism in AIRCoP}
- Pipeline parallelism in AIRCoP is done through module annotation systems. The annotation consists of an integer representing the number of micro-batches (or pipelines).
- Users can split the model into several modules and annotate them for Pipeline parallelism. The resulting training loop is different from naive multiple modules.
+ Pipeline Parallelism in AIRCoP is done through module annotation systems. The annotation consists of an integer representing the number of micro-batches, a.k.a. pipelines.
+ Users can split the model into several modules and annotate them to achieve Pipeline Parallelism. The resulting training loop is different from naively partitioned modules.
  \begin{align*}
@@
 \end{align*}
-Assume we have $\epsilon$ pipelines.
-The target program will have a top-level training loop for iterating mini-batches.
-Inside each mini-batch iteration, there will be a training loop of size $\epsilon$ for iterating the forward passes of the $\epsilon$ micro-batches and a training loop of size $\epsilon$ for iterating the backward passes of the $\epsilon$ micro-batches.
+For $\epsilon$ pipelines, the target program will have a top-level training loop for iterating mini-batches.
+Inside each mini-batch iteration, a training loop of size $\epsilon$ is used to iterate the forward pass of the $\epsilon$ micro-batches, another training loop of size $\epsilon$ is used to iterate the backward pass of the $\epsilon$ micro-batches.
 The gradient will be accumulated inside each backward pass of micro-batches.
-After the backward pass training loop finishes, the weight will be updated based on gradient descent for the entire mini-batch.
+The weight will be updated based on gradient descent for the entire mini-batch when the backward pass training loop ends.
  \section{Formal description of Pipeline Parallelism Transformation}\label{formalpipe}
@@
 %\end{align*}
-Transformation for pipeline parallelism has 2 stages; \textbf{Collection Phase} and \textbf{Generation Phase}.
+Transformation for Pipeline Parallelism has 2 stages: \textbf{Collection Phase} and \textbf{Generation Phase}.
 The Collection Phase collects the necessary information like forward/backward operation.
@@
 During Collection Phase we need to collect the weight operations $E_w$, forward operations $E_f$ and backward operations $E_b$.
-For pipeline parallelism, we will add send/recv into $E_f$ and $E_b$ to maintain the data dependencies between modules.
-The corresponding variables of the operation is $xs^w$, $xs^f$, $xs^b$.
-We also need to calculate the lift variables $$xs^l = Defined(E_f) \bigcap Used(E_b)$$, the lifted variables $xs^l$ are the variables defined in the forward pass and used in the backward pass.
-Because we are generating different loops for forward passes and backward passes of micro-batches, we need to lift $xs^l$ into the outer scope and keep multiple versions of those variables for different micro-batches.\par
+For Pipeline Parallelism, we will add send/recv into $E_f$ and $E_b$ to maintain the data dependencies between modules.
+$xs^w$, $xs^f$, $xs^b$ are the corresponding variables of $E_w$, $E_f$, $E_b$.
+The lift variables are calculated as: $$xs^l = Defined(E_f) \bigcap Used(E_b)$$The lifted variables $xs^l$ are the variables defined in the forward pass and used in the backward pass.
+We need to lift $xs^l$ into the outer scope and keep multiple versions of those variables for different micro-batches because we are generating different loops for forward pass and backward pass of micro-batches.\par
 \paragraph{Generation Phase}
@@
 \end{align*}
-$op_l$ means that the operation will read from the versioned lift variable instead of the original variable.
+where $op_l$ means that the operation will read from the versioned lift variable instead of the original variable.
 $xs^\epsilon = Zero^\epsilon$ will generates an array of $\epsilon$ versions of the variable $xs$.\par
 $G_w$ generates the weight and their gradient and momentum, $G_l$ generate an array of lift nodes for different micro-batches, $G_f$ generates forward pass and write the correct value to the versioned lift variables, $G_b$ generates the backward pass and read from the versioned lift variable, $G_o$ generates the optimization pass.
-Because our source language does not allow mutation, all the mutation operations (Assign/Accumulate/Optimize) are automatically generated in the Generation Phase.
+All the mutation operations (Assign/Accumulate/Optimize) are automatically generated in the Generation Phase because our source language does not allow mutation.
 We only need to write the versioned lift node at the end of the forward pass, and the Generation phase can generate the correct training loop.
@@
 LMS has ANF\cite{flanagan1993essence} form sea of Nodes\cite{click1995simple} IR.
-It also has an effect system for dealing with side effects and dependencies introduced by effects.
-LMS can perform Dead Code Elimination and Common Subexpression Elimination, making it sufficient for our tensor computation graph transformation.
-LMS IR is customizable and offers users the flexibility to define their operation while maintaining the data and effect dependencies between them.
+It also has an effect system to handle side effects and dependencies introduced by effects.
+LMS can perform Dead Code Elimination and Common Subexpression Elimination; therefore is sufficient for implementing our tensor computation graph transformation.
+LMS IR is customizable and offers users the flexibility to define their operation while maintaining the data and effect dependencies.
 We use custom IR nodes to represent Input/Weight Tensor and Machine learning Operations and use the effect system to represent their dependencies.
@@
 The source language does not mutate tensors.
-We use overloading of the typed frontend objects and operations and reflection to generate the corresponding IR of the source program.
+We use overloading of the typed frontend objects and reflection to generate the corresponding IR of the source program.
 We put the explicit TensorType and Annotations into the RHS of the IR to make it persistent and used in later transformation.
@@
 We use the transformation pass of LMS to transform the IR and simulate the tensor graph transformation to generate the correct training loop.
-We first need to perform DimName pass to merge the equivalent dimension of different tensors.
-For example, if matrix multiplication happens between matrixes of dimension $[m, n]$ and dimension $[j,k]$, then dimension $n$ and dimension $j$ can be merged.
-Annotations of tensors and operations represent which dimension to partition, so we have to update those annotations too.
+We first perform DimName pass to merge the equivalent dimension of different tensors, e.g., if matrix multiplication happens between matrixes of dimension $[m, n]$ and dimension $[j,k]$, then dimension $n$ and dimension $j$ can be merged.
+Annotations of tensors and operations represent which dimension to partition, so we must update those annotations.
 \par
@@
 The backward operation is collected in reverse order to maintain the correct semantic of differentiation.
-Because the backward pass should be generated after the forward pass, we collect backward operation in the form of closures.
+We collect backward operations as closures because the backward pass should be generated after the forward pass.
 The Collection Phase will also collect paired send/recv operations at the boundaries of different modules to maintain the correct data dependencies.
 We implement the Collection Phase by defining a function aircopCollect for each IR Node to collect the corresponding information.
-After the Collection Phase, we have the necessary information to perform automatic differentiation (AD).
+After the Collection Phase, we have obtained the necessary information to perform automatic differentiation, i.e., AD.
 We also need to collect lifted tensors to generate the correct training loop.
-As we mentioned before, the forward pass and backward pass of micro-batches are iterated in different loops to achieve pipeline parallelism.
-Therefore, the tensors allocated in the forward pass and used in the backward pass will cause an error because forward and backward passes are now in different scopes.
+As we mentioned before, the forward pass and backward pass of micro-batches are iterated in different loops to achieve Pipeline Parallelism.
+The tensors allocated in the forward pass and used in the backward pass will trigger an error because forward and backward passes are now in different scopes.
 These tensors are called lifted tensors.
@@
 The Generation Phase generates the correct training loop based on the collected information.
-Inside each module, it will first generate the top-level training loop for mini-batches.
-Then it will generate the forward training loop and backward training loop for micro-batches based on Module Annotations and use the collected backward operations to realize automatic differentiation as we mentioned in Section \ref{formalpipe}.
+It will first generate the top-level training loop of mini-batches for each module.
+Inside the body of the top-level loop, as we mentioned in Section \ref{formalpipe}, it will generate the forward training loop and backward training loop for micro-batches based on Module Annotations and use the collected backward operations to realize automatic differentiation.
 Inside both the forward and backward pass, the read and write to the lifted tensors will be transformed into read/write to the versioned tensor array.
@@
 The next pass will perform the Spatial Partition transformation.
-This part is of past work, so we will not dig further into this pass.
-The main purpose of this pass is to partition the tensor and operations based on annotations and add the corresponding inter operation tensor communication for synchronization (AllReduce for MatrixMultiplication or Convolution).\par
+This part is of previous work, so we will not dig further into this pass.
+The main purpose of this pass is to partition the tensors and operations based on annotations and add the corresponding tensor communication between operations for synchronization, e.g., AllReduce for MatrixMultiplication and Convolution.\par
 \subsection{MPI/NCCL Pass}
-The final pass is to generate the corresponding MPI/NCCL instructions for inter-device communications.
-NCCL (NVIDIA Collective Communications Library) is the communication library introduced by NVIDIA for efficient collective and point-to-point communication across multiple GPU and multiple Nodes.
+The final pass generates the corresponding MPI/NCCL instructions for inter-device communications.
+NCCL (NVIDIA Collective Communications Library) is the communication library designed by NVIDIA for efficient collective and point-to-point communication among multiple GPU and multiple Nodes.
 Our target program is CUDA code with MPI/NCCL for communication.
@@
 We will create 2 NCCL communicators, one global communicator for cross-module communication (NCCLSend/Recv) and one local communicator for collective tensor communication (NCCLAllReduce) inside modules.
-We also need to handle tensor communication between modules if they have different Spatial partitions, which can be naively implemented by gather and scatter.
+We also need to handle tensor communication between modules if they have different Spatial partitions, which can be implemented by gather and scatter.
 We use cuBLAS and cuDNN for matrix computation and convolution networks.\par
-We statically allocate GPU memory before the training loop and avoid time-exhausting cudaMalloc/cudaFree inside the training loop to optimize memory management.
-We currently use code-motion techniques to lift the cudaMalloc inside the training loop and pre-allocate the memory before the training loop to optimize memory management.
-This approach could significantly improve performance because cudaMalloc/Free often takes a very long time.\par
-Our generated CUDA code can efficiently represent the correct training loop of machine learning models with both Spatial Partition and Pipeline Parallelism.
+We statically allocate GPU memory before the training loop and avoid time-exhausting cudaMalloc/cudaFree inside the loop to optimize memory management.
+We currently use code-motion techniques to lift the cudaMalloc out of the training loop and pre-allocate the memory to optimize memory management.
+This approach could significantly improve performance because cudaMalloc/Free often occupies excessive time.\par
+Our generated CUDA code can efficiently represent the correct training loop of machine learning models with Spatial Partition and Pipeline Parallelism.
  \section{Evaluation}\label{evaluation}
@@
 The server is a single GPU cluster with a high-speed inter-GPU connection.
-The ideal network topology for Pipeline parallelism is an extensive training system consisting of several Nodes, but we currently could not run the evaluation under that scenario due to limited hardware resources and time.
+The ideal network topology for Pipeline Parallelism is an extensive training system consisting of several Nodes, but we currently could not run the evaluation under that scenario due to limited hardware resources and time.
 The purpose of this benchmark is to validate our graph transformation and gain a deeper insight into the choice of different parallelism strategies.\par
@@
  \begin{itemize}
-   \item Data parallelism running on 2 GPUs
-   \item Data parallelism running on 4 GPUs
-   \item Mixed parallelism running on 4 GPUs --- Naively partitioned into 2 modules with each module running in Data parallelism on 2 GPUs
-   \item Mixed parallelism running on 4 GPUs --- 2 modules in Stacked Pipeline parallelism (4 pipelines) with each module running in Data parallelism on 2 GPUs
-   \item Mixed parallelism running on 4 GPUs --- 2 modules in Stacked Pipeline parallelism (8 pipelines) with each module running in Data parallelism on 2 GPUs
+   \item Data Parallelism running on 2 GPUs
+   \item Data Parallelism running on 4 GPUs
+   \item Mixed Parallelism running on 4 GPUs --- Naively partitioned into 2 modules with each module running in Data Parallelism on 2 GPUs
+   \item Mixed Parallelism running on 4 GPUs --- 2 modules in Stacked Pipeline Parallelism (4 pipelines) with each module running in Data Parallelism on 2 GPUs
+   \item Mixed Parallelism running on 4 GPUs --- 2 modules in Stacked Pipeline Parallelism (8 pipelines) with each module running in Data Parallelism on 2 GPUs
  \end{itemize}
@@
   \includegraphics[scale=0.5]{pipelineparallelruntime.png}
-  \caption{Data Parallelism vs Pipeline parallelism}
+  \caption{Data Parallelism vs Pipeline Parallelism}
   \label{fig:runtime}
@@
   \includegraphics[scale=0.5]{pipelineparallelspeedup.png}
-  \caption{Data Parallelism vs Pipeline parallelism}
+  \caption{Data Parallelism vs Pipeline Parallelism}
   \label{fig:speedup}
@@
 The running time of 5 settings is shown in Figure \ref{fig:runtime}, the corresponding speedup is shown in Figure \ref{fig:speedup}.
-We can notice that simply dividing a model into 2 modules will not cause any speed up due to the low device usage caused by the cross-module dependencies.
+It can be observed that simply dividing a model into 2 modules will not cause any speed up due to the low device usage caused by the cross-module dependencies.
 The device will idle while waiting for data to arrive.
-We can notice that using Pipeline parallelism with 4 pipelines rather than naive module partition on 2 modules will cause a speedup of 1.6x.
-Moreover, we can further accelerate the training loop by increasing the pipeline number (micro-batch number) to 8 and reaching a speedup of 1.77x.
-However, keep in mind that increasing pipeline number will increase memory usage due to versioned tensor.\par
+We can notice that using Pipeline Parallelism with 4 pipelines rather than naive module partition on 2 modules will cause a speedup of 1.6x.
+Moreover, we can further accelerate the training loop by increasing the pipeline number, i.e., micro-batch number, to 8 and reach a speedup of 1.77x.
+However, increasing the pipeline number will cause the memory usage to surge due to versioned tensors.\par
 The speedup achieved is significant and is very close to the theoretical limit.
 No matter how many pipelines we choose, some bubble time will still exist because the different modules of the same micro-batch will still have dependencies between them.
-For 2 modules, the theoretical speedup limit is $\frac{2n}{n+1}$ ($n$ is the pipeline number), which is about the same as our evaluation result.\par
- However, we find out that 2 modules on Stacked Pipeline parallelism with each module running at Data Parallelism on 2 GPUs (Mixed Parallelism) still perform a little worse than Data parallelism on 4 GPUs.
+For 2 modules, the theoretical speedup limit is $\frac{2n}{n+1}$ where $n$ is the pipeline number, which is about the same as our evaluation result.\par
+However, we find that 2 modules on Stacked Pipeline Parallelism with each module running at Data Parallelism on 2 GPUs, i.e., Mixed Parallelism, still perform a little worse than Data Parallelism on 4 GPUs.
 This result is anticipated because we perform our evaluation on a single GPU cluster with very high-speed inter-GPU connections, which makes the frequent tensor communication (ALLReduce) overhead introduced by Data Parallelism negligible.
 However, the cross-module dependencies for micro-batch will still cause some bubble time.
-As we mentioned before, the ideal network topologies for Pipeline parallelism are an extensive training system consisting of multiple GPU clusters connected through the network.
-Then we can use Data parallelism inside each cluster and Pipeline parallelism between clusters.
-We currently could not perform evaluation under this ideal scenario due to limited time and hardware resources, but we could predict that Pipeline parallelism mixed with Data parallelism will perform much better than only using Data parallelism in this scenario because Pipeline parallelism could reduce the communication overhead because inter-cluster communication only happens at module boundaries.\par
+As we mentioned before, the ideal network topologies for Pipeline Parallelism are an extensive training system consisting of multiple GPU clusters connected through the network.
+Then we can use Data Parallelism inside each cluster and Pipeline Parallelism among clusters.
+We currently could not perform evaluation under this ideal scenario due to limited time and hardware resources, but we could deduce that Pipeline Parallelism mixed with Data Parallelism will perform much better than only using Data Parallelism under this scenario because Pipeline Parallelism could significantly reduce the communication overhead by restricting the inter-cluster communication to only happen at module boundaries.\par
  The evaluation shows that our framework is flexible, and the user could switch to different combinations of Parallelism settings to reduce the total communication overhead and achieve optimal performance.
@@
  \section{Conclusion} \label{conclusion}
-In conclusion, we implement Pipeline parallelism for AIRCoP, make it a machine learning framework that can integrate different parallelism strategies, and offer user flexibility to fully explore different combinations of parallel training strategies under different hardware scenarios and gain a deeper insight into parallel training.
+We implement Pipeline Parallelism for AIRCoP, make it a machine learning framework that can integrate different parallelism strategies, and offer user flexibility to fully explore different combinations of parallel training strategies under different hardware scenarios and gain a deeper insight into parallel training.
 \bibliographystyle{acm}

### 2021-12-18 report/scripts/benchmark_plot.py
complete script to generate plot from csv raw data

@@ -1,0 +1,52 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+import matplotlib.ticker as mtick
+
+
+if __name__=="__main__":
+    import pandas as pd
+    import matplotlib.pyplot as plt
+    import seaborn as sns
+
+    # Load data
+    df = pd.read_csv('../data/benchmark.csv', index_col=0)
+
+    # rename index
+    df.rename(index={'Data Parallism (2 GPU)': 'Data Parallism\n(2 GPU)',\
+       'Data Parallism (4 GPU)': 'Data Parallism\n(4 GPU)', 'Mixed Parallism (no pipeline)':\
+       'Mixed Parallism\n(no pipeline)', 'Mixed Parallism (4 pipelines)': 'Mixed Parallism\n(4 pipelines)',\
+       'Mixed Parallism (8 pipelines)': 'Mixed Parallism\n(8 pipelines)'}, inplace=True)
+    df['Runtime'] = df['Runtime'].round(3)
+
+    # plot runtime
+    ax = df.plot.bar(grid=False, figsize=(8,4))
+    plt.xticks(rotation=0)
+    (low, high) = plt.ylim()
+    plt.ylim(low, high*1.1)
+    plt.xlabel("")
+    plt.ylabel("")
+    plt.title("Runtime of 40 iterations (sec)")
+    for col in df.columns:
+      for id, val in enumerate(df[col]):
+        ax.text(id, val+1.4, str(val), va="top", ha="center")
+    plt.legend().remove()
+    plt.savefig("../pictures/pipelineparallelruntime.pdf", format="pdf", bbox_inches="tight")
+
+    # plot speed up
+    dfspeedup = df.rename(columns={'Runtime':'Speed Up'})
+    baseline = dfspeedup.loc['Data Parallism\n(2 GPU)']['Speed Up']
+    dfspeedup['Speed Up'] = baseline / dfspeedup['Speed Up']
+    dfspeedup['Speed Up'] = dfspeedup['Speed Up'].round(2)
+    axspeed = dfspeedup.plot.bar(grid=False, figsize=(8,4))
+    axspeed.yaxis.set_major_formatter(mtick.PercentFormatter(xmax=1, decimals=None, symbol='%', is_latex=False))
+    plt.xticks(rotation=0)
+    (low, high) = plt.ylim()
+    plt.ylim(low, high*1.1)
+    plt.xlabel("")
+    plt.ylabel("")
+    plt.title("Speed Up")
[…]

### 2021-12-18 report/scripts/benchmark_plot.py
fix python script error

@@ -4,9 +4,6 @@ import matplotlib.ticker as mtick
 
 
 if __name__=="__main__":
-    import pandas as pd
-    import matplotlib.pyplot as plt
-    import seaborn as sns
 
     # Load data
     df = pd.read_csv('../data/benchmark.csv', index_col=0)

### 2021-12-19 report/paper.tex
complete script to generate plot from csv raw data
Second round Spelling/Grammar Pass

@@
 Operations are connected through let-bindings.
-Furthermore, we use annotation of the formal syntax to represent different parallelism strategies.
+We use annotation of the formal syntax to represent different parallelism strategies.
 We develop annotation systems for both Tensor, Operation, and Module.
@@
 Stacked Pipeline Parallelism will introduce little communication overhead because it will only transfer data at the module boundaries.
-This feature of Stacked Pipeline Parallelism makes it suitable both for systems with high-speed interconnections and weakly connected systems.
+This feature of Stacked Pipeline Parallelism makes it suitable both for systems with high-speed interconnections as well as weakly connected systems.
 \par
@@
 $G_w$ generates the weight and their gradient and momentum, $G_l$ generate an array of lift nodes for different micro-batches, $G_f$ generates forward pass and write the correct value to the versioned lift variables, $G_b$ generates the backward pass and read from the versioned lift variable, $G_o$ generates the optimization pass.
-All the mutation operations (Assign/Accumulate/Optimize) are automatically generated in the Generation Phase because our source language does not allow mutation.
-We only need to write the versioned lift node at the end of the forward pass, and the Generation phase can generate the correct training loop.
+All the mutation operations (Assign/Accumulate/Optimize) are automatically generated in the Generation Phase because our source language does not allow mutation, so we only need to write the versioned lift nodes at the end of the forward pass.
 \section{Implementation} \label{implementation}
@@
 We use overloading of the typed frontend objects and reflection to generate the corresponding IR of the source program.
-We put the explicit TensorType and Annotations into the RHS of the IR to make it persistent and used in later transformation.
+We put the explicit TensorType and Annotations into the RHS of the IR to make them persistent and use them in later transformation passes.
 \par
@@
 We currently use code-motion techniques to lift the cudaMalloc out of the training loop and pre-allocate the memory to optimize memory management.
-This approach could significantly improve performance because cudaMalloc/Free often occupies excessive time.\par
+This approach could significantly improve performance because cudaMalloc/Free often occupies excessive amount of time.\par
 Our generated CUDA code can efficiently represent the correct training loop of machine learning models with Spatial Partition and Pipeline Parallelism.
@@
   \centering
-  \includegraphics[scale=0.5]{pipelineparallelruntime.png}
+  \includegraphics[scale=0.5]{pipelineparallelruntime.pdf}
   \caption{Data Parallelism vs Pipeline Parallelism}
@@
   \centering
-  \includegraphics[scale=0.5]{pipelineparallelspeedup.png}
+  \includegraphics[scale=0.5]{pipelineparallelspeedup.pdf}
   \caption{Data Parallelism vs Pipeline Parallelism}
@@
 This result is anticipated because we perform our evaluation on a single GPU cluster with very high-speed inter-GPU connections, which makes the frequent tensor communication (ALLReduce) overhead introduced by Data Parallelism negligible.
-However, the cross-module dependencies for micro-batch will still cause some bubble time.
+However, the cross-module dependencies of micro-batches in Pipeline Parallelism will still cause some bubble time.
 As we mentioned before, the ideal network topologies for Pipeline Parallelism are an extensive training system consisting of multiple GPU clusters connected through the network.
